---
title: "SELE_cluster_analysis"
author: "Tim N"
date: "10/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
source("../../global_in/general_scripts.R")
```

```{r}
cluster_list <- AllGeneClusterPuller("../../MARCC output/variance_genes_kendall/")
manual_curation <- read.csv("../../manual_cluster_analysis/manual_curation.csv")
key_dat <- read.csv("../../automated_cluster_analysis/data_out/key_table.csv")
phen_dat <- read.csv("../../global_in/gtex_phenotypes_v8.csv")
samp_dat <- data.table::fread("../../global_in/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt")
death_dat <- read.csv("../../global_in/gtex_COD_phenotype.csv")
```

Find all  clusters with SELE
```{r}
sele_clusters <- row.names(ClusterGeneOverlaps("SELE",cluster_list,only_hits = T))

sele_cluster_df <- as.data.frame(sele_clusters) %>%
    separate(col = sele_clusters, into = c("tissue", "cluster"), sep = "-")
```

Get the Z-scores fort clusters associated with SELE
```{r}
all_profiles <- ZscoreProfilePuller(sele_cluster_df$tissue,"../../MARCC output/variance_profiles_kendall/")
sele_profiles <- NULL
for (row_ind in seq(nrow(sele_cluster_df))) {
    cur_tiss <- sele_cluster_df[row_ind,1]
    cur_clust <- sele_cluster_df[row_ind,2]
    if (cur_clust == "X") {
        next
    }
    temp_dat <- all_profiles[[cur_tiss]][,c("SUBJID", "SAMPID", cur_clust)]
    colnames(temp_dat)[3] <- "SELE_prof"
    temp_dat$tissue <- cur_tiss
    sele_profiles <- rbind(sele_profiles, temp_dat)
}


joined_dat <- sele_profiles %>% left_join(., samp_dat) %>% left_join(., phen_dat) %>% left_join(.,death_dat) %>%
    mutate(DTHHRDY_factor = factor(DTHHRDY, levels = c(1,2,3,4,0)))
```

```{r}
ggplot(joined_dat, aes(SMTSISCH, SELE_prof, color = tissue)) +
    geom_point(alpha = .3) +
    theme(legend.position = "none")

```

```{r}
require(lme4)
require(lmerTest)
ggplot(joined_dat, aes(DTHCODSIM, SELE_prof)) +
    ggforce::geom_sina(alpha = 0.5) +
    theme(legend.position = "none")

initial_result <- lmer(SELE_prof ~ DTHCODSIM * tissue + DTHHRDY_factor +
         SMTSISCH + AGE + (1|SUBJID), data = joined_dat)  %>% broom.mixed::tidy()
```


Complex model
```{r}

model <- lmer(SELE_prof ~ SMTSISCH + DTHHRDY_factor + AGE + SEX + SMGEBTCH + (1|SUBJID), data =joined_dat)
summary(model)
```

Individual tissue SELE
```{r}

all_tissues <- unique(joined_dat$tissue)
model_list <- list()
new_df <- NULL
for (cur_tissue in all_tissues) {
    cur_dat <- filter(joined_dat, tissue == cur_tissue)
    current_model <- lm(SELE_prof ~ SMTSISCH + DTHHRDY_factor + AGE + SEX + SMGEBTCH,
       data = cur_dat)
   cur_summary <- broom::tidy(summary(current_model))
   model_list[[cur_tissue]] <- cur_summary
   ## Add readable output
   terms_we_care <- cur_summary[2:6,]
   terms_we_care$tissue <- cur_tissue
   terms_we_care <- terms_we_care %>% mutate(estimate_direction = ifelse(estimate > 0, "positive", "negative"),
                            sig_o_no = ifelse(p.value < 0.05, "sig", "not_sig"))
   new_df <- rbind(new_df, terms_we_care)
}

table(new_df$term, new_df$sig_o_no)

hist(new_df$p.value)
```
How often is PLAG2A and SELE together or opposite of each other?
```{r}
sele_vs_plag <- GenesFromTissues("", c("PLA2G2A", "SELE"), key_dat, "../../Setup Code/output/", TRUE)
cor.test(sele_vs_plag$SELE, sele_vs_plag$PLA2G2A,method = "kendall")
```

```{r}
sele_and_plag <- ClusterGeneOverlaps(c("SELE", "PLA2G2A"),cluster_list,only_hits = T)
```


See  how different tissues deal with MI vs Anoxia
```{r}

## Filter data to MI and anosmia
mi_anox <- filter(joined_dat, DTHCODSIM %in% c("MI", "Anoxia"))
## Generate random tissues
rand_tisses <- sample(unique(mi_anox$tissue), replace = F,size = 4)

temp_test <- filter(mi_anox, tissue %in% rand_tisses)


ggplot(temp_test, aes(tissue, SELE_prof, color = DTHCODSIM)) +
    ggforce::geom_sina() + coord_flip()

ggplot(mi_anox, aes(tissue, SELE_prof, color = DTHCODSIM)) +
    ggforce::geom_sina() + coord_flip()

ggsave("data_out/anoxia_v_MI_sele.pdf", width = 8, height = 24)
```

Test if IL6, SELE, and TGFB1 are expressed in MI
```{r}

mi_gene_test <- GenesFromTissues(tiss_of_int = c(""),
                                    genes_of_int = c("IL6", "SELE", "TGFB1","KLF2", "PECAM1"),
                                                     key_dat =  key_dat,
                                                        "../../Setup Code/output/", T)

mi_gene_test_wide <- GenesFromTissWiden(mi_gene_test)
mi_gene_test_heart_only <- mi_gene_test %>% filter(startsWith(tissue, "heart")) %>%
    left_join(., death_dat) %>%
    left_join(., phen_dat) %>%
    left_join(., samp_dat) %>%
    mutate(MI_or_not = ifelse(DTHCODSIM == "MI", "MI", "not_MI"),
           MI_or_not = factor(MI_or_not, levels = c("not_MI", "MI")))

mi_gene_test <- inner_join(mi_gene_test, mi_anox)


ggplot(mi_gene_test, aes(tissue, KLF2, color = DTHCODSIM)) +
    ggforce::geom_sina() +
    coord_flip()


ggplot(mi_gene_test, aes(SELE, KLF2, color = tissue)) +
    geom_point(alpha = 0.3) +
    theme(legend.position = "none")

ggplot(mi_gene_test, aes(SELE, PECAM1, color = tissue)) +
    geom_point(alpha = 0.3) +
    theme(legend.position = "none")

ggplot(mi_gene_test, aes(KLF2, PECAM1, color = tissue)) +
    geom_point(alpha = 0.3) +
    theme(legend.position = "none")



    

ggplot(mi_gene_test_heart_only, aes(MI_or_not, IL6)) +
    ggforce::geom_sina() +
    facet_grid(~tissue)

ggplot(mi_gene_test_heart_only, aes(MI_or_not, SELE)) +
    ggforce::geom_sina() +
    facet_grid(~tissue)

ggplot(mi_gene_test_heart_only, aes(IL6, SELE, color = tissue, shape = MI_or_not)) +
    geom_point()

summary(lm(IL6 ~MI_or_not * tissue + AGE+ SEX+ SMTSISCH + factor(DTHHRDY), data = mi_gene_test_heart_only))

ggplot(mi_gene_test_heart_only, aes(SMTSISCH, IL6)) +
    geom_point()

ggplot(mi_gene_test_heart_only, aes(MI_or_not, SELE)) +
    ggforce::geom_sina()

## Separate heart checks
hearts <- unique(mi_gene_test_heart_only$tissue)
heart_results <- list()
for (cur_heart in hearts) {
    ind <- mi_gene_test_heart_only$tissue == cur_heart
    pre_test <- mi_gene_test_heart_only[ind,]
    cur_model_IL6 <- lm(IL6 ~ MI_or_not + AGE + SEX + SMTSISCH + factor(DTHHRDY),
                    data = pre_test) %>% summary()
    cur_model_SELE <- lm(SELE ~ MI_or_not + AGE + SEX + SMTSISCH + factor(DTHHRDY),
                    data = pre_test) %>% summary()
    heart_results[[cur_heart]][["IL6"]] <- cur_model_IL6
    heart_results[[cur_heart]][["SELE"]] <- cur_model_SELE
}
```

Brain only analysis
```{r}
brain_dat  <- filter(joined_dat, startsWith(tissue, "brain"))
no_brain_dat <- filter(joined_dat, !startsWith(tissue, "brain"))
ggplot(brain_dat, aes(DTHCODSIM, SELE_prof)) +
    ggforce::geom_sina(alpha = 0.5) +
    theme(legend.position = "none")

ggplot(no_brain_dat, aes(DTHCODSIM, SELE_prof)) +
    ggforce::geom_sina(alpha = 0.5) +
    theme(legend.position = "none")

brain_tf_dat <- mutate(joined_dat, brain_tiss = ifelse(startsWith(tissue, "brain"), "brain_tiss", "not_brain"))

ggplot(brain_tf_dat, aes(DTHCODSIM, SELE_prof)) +
    ggforce::geom_sina(alpha = 0.5) +
    theme(legend.position = "none") + facet_grid(~brain_tiss) +
    theme(axis.text.x = element_text(angle = 80, hjust=1))
    
```

Myocardial infarction is a an artefact
```{r}

no_brain_dat <- mutate(no_brain_dat,
                       MI_or_not = ifelse(DTHCODSIM == "MI", "MI", "not_MI"))
all_no_brain_tissues <- unique(no_brain_dat$tissue)
pdf("data_out/MI_artefact.pdf", height = 6, width = 10)
for (cur_tiss_ind in seq_along(all_no_brain_tissues)) {
    cur_tiss <- all_no_brain_tissues[[cur_tiss_ind]]
    ind <- no_brain_dat$tissue == cur_tiss
    cur_dat <- no_brain_dat[ind,]
    plt <- ggplot(cur_dat, aes(DTHHRDY_factor, SELE_prof, color = MI_or_not)) +
        ggforce::geom_sina() +
        labs(title = paste("Tissue:", cur_tiss))
    print(plt)
}
dev.off()



pdf("data_out/MI_artefact_4weeks.pdf", height = 6, width = 10)
for (cur_tiss_ind in seq_along(all_no_brain_tissues)) {
    cur_tiss <- all_no_brain_tissues[[cur_tiss_ind]]
    ind <- no_brain_dat$tissue == cur_tiss
    cur_dat <- no_brain_dat[ind,]
    plt <- ggplot(cur_dat, aes(DTHHRDY_factor, SELE_prof, color = factor(MHNPHYS4W))) +
        ggforce::geom_sina() +
        labs(title = paste("Tissue:", cur_tiss))
    print(plt)
}
dev.off()
```




SELE obesity test
```{r}

fat_only <- filter(joined_dat, startsWith(tissue, "adipose"))

ggplot(fat_only, aes(BMI, SELE_prof, color = tissue)) +
    geom_point()
```

Time after stroke?

```{r}
stroke_only <- brain_dat %>% filter(DTHCODSIM == "Stroke") 

ggplot(stroke_only, aes(TRISCHD, SELE_prof)) +
    geom_point()

```
Shared hits plot
```{r}
SharedHits <- function(gene_of_int, cluster_list, min_appearances,
                       cluster_rows = T, cluster_cols = T,
                       gene_text_size = 7){
    ### The purpose of this function is to find out what tissues are sharing
    ### the most overlapping genes in a cluster based on another gene. e.g.
    ### what tissues have the most similar clusters of tissues that have SELE?
    
    ### Get a total vector of genes that cluster at any point with SELE
    clust_int_list <- list()
    for (tiss_ind in seq_along(cluster_list)) {
        cur_tiss <- cluster_list[[tiss_ind]]
        cur_tiss_name <- names(cluster_list)[[tiss_ind]]
        int_find <- unlist(lapply(cur_tiss, function(x){gene_of_int %in% x}))
        if (!(TRUE %in% int_find)) {
            next
        }
        if (names(int_find)[int_find==TRUE] == "X") {
            next
        }
        clust_int_list[[cur_tiss_name]] <- unname(unlist(cur_tiss[int_find]))
    }
    ### Possibly filter as a table to the min_appearances
    gene_overlap_table <- sort(table(unlist(clust_int_list)),decreasing = T)
    most_common_genes <- names(gene_overlap_table)[gene_overlap_table >= min_appearances]
    ### create a matrix with the columns being tissues and the y axis being
    ### gene overlaps
    new_matrix <- matrix(nrow = length(most_common_genes),
                         ncol = length(clust_int_list))
    colnames(new_matrix) <- names(clust_int_list)
    rownames(new_matrix) <- most_common_genes
    
    ## Might be a way to do this with apply, but rather do it right than fast
    for (cur_ind in seq(ncol(new_matrix))) {
        cur_name <- colnames(new_matrix)[cur_ind]
        temp_result <- rownames(new_matrix) %in%  clust_int_list[[cur_name]]
        temp_result <- as.numeric(temp_result)
        new_matrix[,cur_ind] <- temp_result
    }
    
    ### Make pheatmap too
    map_out <- pheatmap(new_matrix,
                        cluster_rows = cluster_rows,
                        cluster_cols = cluster_cols,
                        fontsize_row = gene_text_size,
                        border_color = "black")
    ### Wrap together into a single output
    final_list <- list()
    final_list[["heatmap_res"]] <- map_out
    final_list[["matrix_res"]] <- new_matrix
    return(final_list)
}

SELE_share <- SharedHits("SELE", cluster_list, min_appearances = 2,
                         gene_text_size = 5)
print(SELE_share$heatmap_res)
pdf(file = "data_out/SELE_overlaps_prototype.pdf", height = 20, width = 10)
print(SELE_share$heatmap_res)
dev.off()



PLA2G2A_share <- SharedHits("PLA2G2A", cluster_list, min_appearances = 2)
print(PLA2G2A_share$heatmap_res)
pdf(file = "data_out/PLA2G2A_overlaps_prototype.pdf", height = 20, width = 10)
print(PLA2G2A_share$heatmap_res)
dev.off()
```

