---
title: "Artery comparisons"
author: "Tim Nieuwenhuis"
date: "7/13/2021"
output: html_document
---

GTEx has three artery tissue, coronary, tibial, and aorta. We would expect some
simialir clusters and some different ones (muscle contamination), we want to see
what the overlap looks like. We will explore this question here.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pheatmap)
source("../../global_in/general_scripts.R")
```


Load in the files
```{r}
tib_genes <- read.csv("../../MARCC output/variance_genes_kendall/kendall-artery_tibial-gene-clusters-high-variance.csv")
cora_genes <- read.csv("../../MARCC output/variance_genes_kendall/kendall-artery_coronary-gene-clusters-high-variance.csv")
aorta_genes <- read.csv("../../MARCC output/variance_genes_kendall/kendall-artery_aorta-gene-clusters-high-variance.csv") 
artery_df_list <- list("tibial" = tib_genes, "coronary" = cora_genes, "aorta"= aorta_genes)
```

Turn tables into lists
```{r}

TableListToGeneList <- function(table_list){
    out_list <- list()
    for (ind_table in seq_along(table_list)) {
        cur_name <- names(table_list)[ind_table]
        cur_df <- table_list[[ind_table]]
        clust_list <- list()
        for (ind_clust in seq(ncol(cur_df))) {
            clust_name <- names(cur_df)[ind_clust]
            clust_list[[clust_name]] <- cur_df[[ind_clust]][cur_df[[ind_clust]] != ""]
        }
        out_list[[cur_name]] <-clust_list
    }
    return(out_list)
}

artery_gene_list <- TableListToGeneList(artery_df_list)

```
Generate overlap pseudo-correlation matrix
```{r}

gene_list <- artery_gene_list

PseudoCorrelate <- function(gene_list, proportion_scale = FALSE){
    n_clusts <- sum(unlist(lapply(gene_list, length)))
    temp_clusts <- lapply(gene_list, names)
    col_row_names <- c()
    for (ind in seq_along(temp_clusts)) {
        cur_res <-paste(names(temp_clusts)[ind], temp_clusts[[ind]],sep = ".")
        col_row_names <- append(col_row_names, cur_res)
    }
    pseudo_matrix <- matrix(data = NA,
                            nrow = n_clusts, ncol = n_clusts)
    colnames(pseudo_matrix) <- col_row_names
    rownames(pseudo_matrix) <- col_row_names
    split_names <- str_split(col_row_names, "[.]", n = 2)
    tiss_names <- sapply(split_names, function(x){x[1]})
    clust_names <- sapply(split_names, function(x){x[2]})
    for (ind_col in seq(ncol(pseudo_matrix))) {
        col_name <- colnames(pseudo_matrix)[ind_col]
        for (ind_row in seq(nrow(pseudo_matrix))) {
            if (!is.na(pseudo_matrix[ind_row, ind_col])) {
                next 
            }
            col_genes <- gene_list[[tiss_names[ind_col]]][[clust_names[ind_col]]]
            row_genes <- gene_list[[tiss_names[ind_row]]][[clust_names[ind_row]]]
            overlap <- sum(col_genes %in% row_genes)
            if (proportion_scale == TRUE) {
               overlap <- overlap/max(c(length(col_genes), length(row_genes)))    
            }
            pseudo_matrix[ind_row, ind_col] <- overlap
            if (ind_row != ind_col) {
                pseudo_matrix[ind_col, ind_row] <- overlap
            }
        }
    }
    return(pseudo_matrix)
}


artery_matrix <- PseudoCorrelate(gene_list = artery_gene_list,proportion_scale = T)

## Create diagnal matrix
length_of_clusts <- unlist(lapply(artery_gene_list, function(x){lapply(x, length)}))
for_plotter <- matrix(data = "", nrow = length_of_clusts, ncol = length_of_clusts)
for (ind in seq_along(length_of_clusts)) {
    for_plotter[ind,ind] <- length_of_clusts[ind]
}

pheatmap(artery_matrix,width = 10, height = 8,filename = "data_out/pseudo_correlations.pdf",display_numbers = for_plotter)
```


See if cluster correlate with each other in the same people
```{r}
arteries <- c("aorta", "tibial", "coronary")
profile_list <- list()
for (artery_type in arteries) {
    cur_art <- read.csv(paste0("../../MARCC output/variance_profiles_kendall/kendall-artery_",
                                artery_type,"-cluster-profiles.csv"))
    colnames(cur_art)[1] <- "SAMPID"
    colnames(cur_art) <- paste(colnames(cur_art), artery_type,sep = "_")
    cur_art$SUBJID <- GTEx_SAMPID_to_SUBJID(cur_art$SAMPID)
    profile_list[[artery_type]] <- cur_art
}

aort_tib <- inner_join(profile_list$aorta, profile_list$tibial, suffix = c("_aorta", "_tibial"), by = "SUBJID")
all_profiles <- inner_join(aort_tib, profile_list$coronary)



all_numeric_profiles <- select(all_profiles, -starts_with("SAMPID"), -SUBJID)

pearson_correlation <- cor(all_numeric_profiles)

pheatmap(pearson_correlation,filename = "data_out/all_profile_correlations.pdf")

```

```{r}
pheno_profiles <- left_join(all_profiles, pheno_dat)

ggplot(pheno_profiles, aes(factor(MHBCTINF), A.1_aorta)) +
    ggforce::geom_sina()

ggplot(pheno_profiles, aes(C.1_tibial, A.1_aorta)) +
    geom_point()
```

