---
title: "base_Z_score_questions.rmd"
author: "Tim Nieuwenhuis"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
source("../../global_in/general_scripts.R")
```

Select items of interest
```{r}
## Tissue R name
tiss <- "muscle_skeletal"
```

load in data
```{r}

#Load genes and profiles
gene_path <- paste0("../../MARCC output/variance_genes_kendall/kendall-",
                    tiss,"-gene-clusters-high-variance.csv")

clust_genes <- read.csv(gene_path)
prof_path <- paste0("../../MARCC output/variance_profiles_kendall/kendall-",
                    tiss,"-cluster-profiles.csv")
clust_profs <- read.csv(prof_path,row.names = 1)

## Load gene clusters

## Load SUBJ data
phen_dat <- read.csv("../../global_in/gtex_phenotypes_v8.csv")
## Load samp data.
samp_dat <- read.delim("../../global_in/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt", fill=T,
                      stringsAsFactors=FALSE, sep="\t", header=T)
## Load common subj_dat
subj_dat <- read.delim("../../global_in/GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt", fill=T,
                      stringsAsFactors=FALSE, sep="\t", header=T)
## Load keys 
key_dat <- read.csv("../../automated_cluster_analysis/data_out/key_table.csv")
```


## Load sample data

Clean and join data


Join_data
```{r}
final_dat <- clust_profs %>% rownames_to_column(var = "SAMPID") %>%
    mutate(SUBJID = GTEx_SAMPID_to_SUBJID(SAMPID)) %>%
    left_join(phen_dat, by = "SUBJID") %>%
    left_join(samp_dat, by = "SAMPID") 

```

Ask questions below
```{r}
ggplot(final_dat, aes(x = SMTSISCH, y = G, color = as.factor(DTHVNT))) +
    geom_point(alpha = .5)

lm_dat <- filter(final_dat, DTHVNT != 99, !is.na(DTHVNT))

summary(lm(G ~ SMTSISCH + DTHVNT + AGE + SEX, data = lm_dat))


ggplot(final_dat, aes(x = SMTSISCH, y = E.2, color = as.factor(DTHVNT))) +
    geom_point(alpha = .5)

ggplot(final_dat, aes(x = SMRIN, y = E.2, color = as.factor(DTHHRDY))) +
    geom_point(alpha = .5)

ggplot(final_dat, aes(x = SMTSISCH, y = E.2, color = as.factor(DTHHRDY))) +
    geom_point(alpha = .5)

ggplot(final_dat, aes(x = SMTSISCH, y = E.1, color = as.factor(DTHHRDY))) +
    geom_point(alpha = .5)

ggplot(final_dat, aes(x = SMTSISCH, y = D, color = as.factor(DTHHRDY))) +
    geom_point(alpha = .5)



```
Finding overlaps between muscle genes and fast and slow twitch genes in our
dataset
```{r}
mouse_twitch_dat <- read.csv("data_in/mouse_skeletal_fast_slow_markers.csv")
mouse_twitch_dat <- mutate(mouse_twitch_dat,
                           twitch = ifelse(average_logFC_scRNA.seq > 0, "fast", "slow"),
                           human_gene = toupper(Gene))
all_clusters<- AllGeneClusterPuller("../../MARCC output/variance_genes_kendall/")
muscle_cluster <- all_clusters$muscle_skeletal

fast_twitch <- filter(mouse_twitch_dat, twitch == "fast")$human_gene
slow_twitch <- filter(mouse_twitch_dat, twitch == "slow")$human_gene

unlist(lapply(muscle_cluster, function(x){ sum(x %in% fast_twitch)}))
unlist(lapply(muscle_cluster, function(x){ sum(x %in% slow_twitch)}))

## No clear fast vs slow twitch, maybe E.2 has more fast twitch and A has more
## slow twitch but that's about it. Let me find out which genes come from which
## clusters though

fast_col <- t(as.data.frame(lapply(
    muscle_cluster, function(x){ paste(x[x %in% fast_twitch],collapse = "; ")})))
slow_col <- t(as.data.frame(lapply(
    muscle_cluster, function(x){ paste(x[x %in% slow_twitch],collapse = "; ")})))
twitch_comparison <- cbind(fast_col,slow_col)
colnames(twitch_comparison) <- c("fast_twitch", "slow_twitch")
```
Limit data to only ventillator individuals to test the "active before death" hypothesis

```{r}

vent_final <- final_dat %>%
    filter(DTHVNT == 1) %>%
    mutate(vent_group = case_when(
        DTHVNTD >= 24 & TRISCHD > 500 ~ "group_2",
        DTHVNTD <24 & TRISCHD > 500 ~ "group_1",
        TRUE ~ "group_3"
    ))



ggplot(vent_final, aes(DTHVNTD, TRISCHD, color = factor(vent_group))) +
    geom_point() +
    labs(title = "Time on Vent Compared to Ischemic Time",
         x = "Log10 Hours on Ventillator",
         y = "Ischemic Time") +
    scale_x_continuous(
        trans = "log10",
        breaks = c(.5, 1, 5, 10, 24, 100, 300, 1000, 10000)
    )


ggplot(vent_final, aes(x = SMTSISCH, y = E.2, color = as.factor(vent_group))) +
    geom_point(alpha = .5)

ggplot(final_dat, aes(x =  as.factor(DTHHRDY), y = E.2, color = as.factor(DTHHRDY))) +
    ggforce::geom_sina(alpha = .5)

ggplot(vent_final, aes(x =  as.factor(vent_group), y = E.2, color = as.factor(vent_group))) +
    ggforce::geom_sina(alpha = .5)



ggplot(final_dat, aes(x =  as.factor(MHNPHYS4W), y = E.2, color = as.factor(MHNPHYS4W), shape = as.factor(DTHHRDY))) +
    ggforce::geom_sina(alpha = .5)


```

Histology randomizer
```{r}
top_muscle <- head(arrange(final_dat, desc(E.2)), 5)
bot_muscle <- head(arrange(final_dat, (E.2)), 5)

pre_shuffle <- rbind(top_muscle, bot_muscle) %>% select(SAMPID) %>%
    mutate(Marc_notes = "")
set.seed(123)
shuffler <- sample(nrow(pre_shuffle))
shuffled <- pre_shuffle[shuffler,]
write.csv(shuffled,"data_out/blinded_muscle_samples.csv",row.names = F)

shuffled$SAMPID %in% top_muscle$SAMPID


marc_labels <- read.csv(file = "data_in/blinded_muscle_samples_marc_labeled.csv")

marc_labels <- left_join(marc_labels, final_dat)


ggplot(marc_labels, aes(Marc_notes, E.2)) +
    ggforce::geom_sina()

```

Fat atrophy checker

```{r}
final_dat <- final_dat %>%
    mutate(fat_mention = case_when(
    str_detect(SMPTHNTS, "fat") ~ TRUE,
    TRUE ~ FALSE),
    atrophy_mention = case_when(
        str_detect(SMPTHNTS, "atroph") ~ TRUE,
                   TRUE ~ FALSE)
    )
)


ggplot(final_dat, aes(x =  atrophy_mention, y = E.2, color =atrophy_mention)) +
    ggforce::geom_sina(alpha = .5)


ggplot(final_dat, aes(x =  fat_mention, y = E.2, color =fat_mention)) +
    ggforce::geom_sina(alpha = .5)


test_out <- CorrectionPlotter(final_dat, "E.2", c("AGE", "SEX"),"SMTSISCH")

test_out$plot + geom_point(aes(color = as.factor(DTHHRDY)))
```

Compare our results with people who have critical illness myopathy

```{r}
sig_CIM_genes <- data.table::fread("https://static-content.springer.com/esm/art%3A10.1186%2Fs13395-019-0194-1/MediaObjects/13395_2019_194_MOESM2_ESM.txt") %>%
    filter(FDR < 0.05)


sig_CIM_genes <- sig_CIM_genes %>%
    mutate(direction = ifelse(`FoldChange(ICU Patients vs. Control)` > 0,
                              "higher_in_ICU",
                              "lower_in_ICU"))

high_ICU_genes <- sig_CIM_genes$`Column ID`[sig_CIM_genes$direction =="higher_in_ICU"]
low_ICU_genes <- sig_CIM_genes$`Column ID`[sig_CIM_genes$direction =="lower_in_ICU"]


high_icu_col <- t(as.data.frame(lapply(
    muscle_cluster, function(x){ paste(x[x %in% high_ICU_genes],collapse = "; ")})))
low_icu_col <- t(as.data.frame(lapply(
    muscle_cluster, function(x){ paste(x[x %in% low_ICU_genes],collapse = "; ")})))
icu_comparison <- as.data.frame(cbind(high_icu_col,low_icu_col))
icu_comparison

TotalOverlapFunction <- function(x, icu_df){sum(x %in% icu_df)}
AllGenesInCluster <- function(x){length(x)}
overlap_high_icu <- lapply(muscle_cluster,  function(x, icu_df){sum(x %in% high_ICU_genes)})
overlap_low_icu <- lapply(muscle_cluster,  function(x, icu_df){sum(x %in% low_ICU_genes)})
total_genes_in_clust <- lapply(muscle_cluster, function(x){length(x)})

pre_bind <-as.data.frame(cbind(unlist(total_genes_in_clust),
                               unlist(overlap_high_icu),
                               unlist(overlap_low_icu)))

new_cim_table <- cbind(pre_bind,icu_comparison)
colnames(new_cim_table) <- c("N of genes", "N overlap with MIC upreg", "N overlap MIC downreg",
                             "Genes overlap MIC upreg", "Genes overlap MIC downreg")
#TNN1 downregulate in MIC according to paper, meaning lower levels are associated with MIC

write.csv(new_cim_table, "data_out/CIM_overlaps.csv")


```

```{r}
ggplot(final_dat, aes(factor(DTHHRDY), J, color = factor(DTHHRDY))) +
    ggforce::geom_sina()
```

