---
title: "Where do genes show up tool"
author: "Tim Nieuwenhuis"
date: "6/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
source("../../global_in/general_scripts.R")
```


```{r}
key_dat <- read.csv("../../automated_cluster_analysis/data_out/key_table.csv")
list_dat <- AllGeneClusterPuller("../../MARCC output/variance_genes_kendall/",
                                 x_clust = FALSE)
```


This was the previous analysis I'm going to do a newer one below more relevant
to the current  question and includes both B.1 and B.2

Genes of interest
```{r}

clusts <- read.csv("../../MARCC output/variance_genes_kendall/kendall-whole_blood-gene-clusters-high-variance.csv")
genes_o_int <- clusts[clusts[,"B.2"] != "", "B.2"]
```

Get genes most commonly shared
```{r}
gene_lapply_out <-  lapply(genes_o_int, function(x){
        res_dat <- lapply(list_dat, function(y){
           ind<- lapply(y, function(z){
                x %in% z
            })
           y[unlist(ind)]
        })
        temp_df <- as.data.frame(table(unlist(res_dat)))
        colnames(temp_df) <- c("genes", x)
        temp_df
    })

for (df_ind in seq_along(gene_lapply_out)) {
    if (df_ind == 1) {
        df_out <- gene_lapply_out[[df_ind]]
    } else{
        df_out <- full_join(df_out, gene_lapply_out[[df_ind]], by = "genes")
    }
}

df_out$most_shared <- column_to_rownames(df_out, "genes") %>% rowSums()

arrange(df_out, desc(most_shared))
```

Get clusters for highest enrichment of gene list
```{r}
cluster_enrichment<- lapply(list_dat, function(x){
    first_lap <- lapply(x, function(y){
        sum(genes_o_int %in% y)
    })
    unlist(first_lap)
})

top_clusters <- sort(unlist(cluster_enrichment),decreasing = T)

top_clusters_filt <- as.data.frame(top_clusters) %>% filter(top_clusters >0)
top_clusters_filt

```
Get the genes from the top clusters
```{r}
clusters_out <- NULL
for (cluster in rownames(head(top_clusters_filt,5))) {
    parts <- unlist(str_split(cluster, "[.]"))
    tissue <- parts[1]
    if (length(parts) == 3) {
        cluster <- paste(parts[2], parts[3],sep = ".")
    } else{
         cluster <- parts[2]
    }
    cur_genes <- list_dat[[tissue]][[cluster]]
    clusters_out[[paste(tissue, cluster, sep = ".")]] <- cur_genes
}
clusters_out
```

New Analysis begins here:

We want to find for both B.1 and B.2 what the top overlapping genes are with
other tissues to see if blood specific ischemia is related to other clusters.


```{r}
## Pull out genes as list to use lapply for analyses
b_clust_genes <- list_dat$whole_blood[c("B.1", "B.2")]

ClusterGeneOverlaps(b_clust_genes$B.1, list_dat)
b_clust_overlaps <- lapply(b_clust_genes, function(x){
    ClusterGeneOverlaps(x, list_dat) %>%
        arrange(desc(n_intersect))
})

most_frequently_found_genes <- lapply(b_clust_overlaps, function(x){
    sort(table(unlist(str_split(x$found_genes,pattern =  "; "))),decreasing = T)
})

most_frequently_found_genes$B.2
```

The above findings create two situations which I would like to examine further
for each cluster. For B.1 I want to know if cluster containing CD163 are often
times correlating subclusters with hemoglobin genes. For B.2 I want to see if 
higher levels of these genes are global throughout the body.


Do CD163 clusters frequently correlate with HBA clusters?
```{r}
gene_1 <- "CD163"
gene_2 <- "HBA1"
cluster_list <- list_dat
profile_loc_or_frame <- "../../MARCC output/variance_profiles_kendall/"

DoGeneSubclustersCorrelate <- function(gene_1,
                                       gene_2,
                                       cluster_list,
                                       profile_loc_or_frame){
    
    ### Output will be a table with the following columns
    ### Tissue of interest
    ### The pearson correlation between clusters with both genes (NA if not both)
    ### Gene_1 what cluster is it in
    ### gene_2 what cluster is it in
    
    ## Find tissues where genes are found in separate clusters or only in one tissue
    table_return <- NULL
    tiss_vect_for_profile <- c()
    for (tiss_ind  in seq_along(cluster_list)) {
        cur_tiss <- cluster_list[[tiss_ind]]
        tiss_name <- names(cluster_list)[tiss_ind]
        gene_vect <- c(gene_1,gene_2)
        overlap_ind <- gene_vect %in% unlist(cur_tiss)
        overlap_val <- sum(overlap_ind)
        ## If there are no overlapping genes skip
        if (overlap_val == 0) {
            next
        } else if (overlap_val == 1){
            ## If there is one overlapping gene record the cluster it is in
            ## but have the correlation be an NA
            found_gene <- gene_vect[overlap_ind]
            bool_ind <-lapply(cur_tiss, function(x){
                found_gene %in% x
            })
            bool_ind <- unlist(bool_ind)
            cluster_found <- names(cur_tiss)[bool_ind]
            overlap_pre_bind <- overlap_ind
            overlap_pre_bind[overlap_ind == TRUE] <- cluster_found
            new_row <- cbind(tiss_name, NA, overlap_pre_bind[1], overlap_pre_bind[2])
            table_return <- rbind(table_return, new_row)
        } else {
            ## Check if they're in the same cluster
            bool_ind <-lapply(cur_tiss, function(x){
                sum(gene_vect %in% x) > 1
            })
            ## If they are in the same cluster add to the table
            if (TRUE %in% unlist(bool_ind)) {
                genes_clust <- names(cur_tiss)[unlist(bool_ind)]
                new_row <- cbind(tiss_name, 1, genes_clust, genes_clust)
                table_return <- rbind(table_return, new_row)
            } else {
                tiss_vect_for_profile <- append(tiss_vect_for_profile, tiss_name)
            }
            
        }
    }
    ## Now pull profiles and filter them for relevant clusters
    if (is.character(profile_loc_or_frame)) {
        profile_loc_or_frame <- ZscoreProfilePuller(tiss_vect_for_profile,
                                                    profile_loc_or_frame)
    } else{
        profile_loc_or_frame_ind <- names(profile_loc_or_frame) %in% tiss_vect_for_profile
        profile_loc_or_frame <- profile_loc_or_frame[profile_loc_or_frame_ind]
    }
    ## Loop over profiles and correlate them
    for (cur_ind in seq_along(profile_loc_or_frame)) {
        cur_prof <- profile_loc_or_frame[[cur_ind]]
        cur_tiss <- names(profile_loc_or_frame)[cur_ind]
        gene_finder <- lapply(cluster_list[[cur_tiss]], function(x){
            sum(gene_vect %in% x)
        })
        clusters_to_corr <- names(gene_finder)[unlist(gene_finder) == 1]
        cor_result <- cor.test(cur_prof[[clusters_to_corr[1]]],
                               cur_prof[[clusters_to_corr[2]]],
                 method = "pearson")
        ## Make sure the tissues are in the right order
        gene_finder_2 <- lapply(cluster_list[[cur_tiss]], function(x){
            gene_vect[gene_vect %in% x]
        })
        gene_finder_2 <- unlist(gene_finder_2)[length( unlist(gene_finder_2)) > 0]
        overlap_pre_bind <- names(gene_finder_2)[match(gene_vect,gene_finder_2)]
        new_row <- cbind(cur_tiss, round(cor_result$estimate,2),
                         overlap_pre_bind[1], overlap_pre_bind[2])
        table_return <- rbind(table_return, new_row)
    }
    table_return <- as_tibble(table_return)
    colnames(table_return) <- c("tissue", "correlation", gene_vect)
    table_return$correlation <- as.numeric(table_return$correlation)
    table_return <- arrange(table_return, desc(abs(correlation)))
    return(table_return)
                                       }

b1_carryover_test <- DoGeneSubclustersCorrelate("CD163", "HBA1", list_dat, "../../MARCC output/variance_profiles_kendall/")
```

So blood invasion (as measured by HBA1) does not positively correlate with CD163.


Now we will look at the top genes across tissues from cluster B.2
```{r}
b2_genes <- names(most_frequently_found_genes$B.2)[most_frequently_found_genes$B.2 > 10]

across_tissue_out <- GenesFromTissues(genes_of_int = b2_genes, key_dat = key_dat,
                 vst_location = "../../Setup Code/output/",
                 run_all_available = T)
## Make inter tissue z-scores
NewZScoreMaker <- function(gene_from_tissue_long_out){
    numeric_ind <- sapply(gene_from_tissue_long_out, is.numeric)
    zscores = rep(NA, nrow(gene_from_tissue_long_out))
    for (cur_tiss in unique(gene_from_tissue_long_out$tissue)) {
        tiss_ind <- gene_from_tissue_long_out$tissue %in% cur_tiss
        numeric_matrix <- gene_from_tissue_long_out[tiss_ind,numeric_ind]
        centered_matrix <- sweep(numeric_matrix, colMeans(numeric_matrix), MARGIN = 2)
        zscore_matrix <- sweep(centered_matrix,
                               apply(numeric_matrix, 2, sd),
                               FUN = "/",
                               MARGIN = 2)
        zscores[tiss_ind] <- apply(zscore_matrix, 1, sum, na.rm =T)/ncol(zscore_matrix)
    }
    return(zscores)    
    
}

across_tissue_out$zscores <- NewZScoreMaker(across_tissue_out)




```

Quick correlation peek

```{r}
temp_comp <- across_tissue_out %>% filter(tissue == sample(unique(across_tissue_out$tissue),size = 2))

GenesFromTissWiden(temp_comp) %>%
    ggplot(., aes_string(x = paste0("zscores_", unique(temp_comp$tissue)[1]),
                         y = paste0("zscores_", unique(temp_comp$tissue)[2]))) +
    geom_point()
```
Take top and bottom whole blood individuals
```{r}
n_of_samps <- 10
ordered_blood <- across_tissue_out %>%
    filter(tissue == "whole_blood") %>%
    arrange(desc(zscores))
top_subj <- head(ordered_blood$SUBJID, n_of_samps)
bot_subj <- tail(ordered_blood$SUBJID, n_of_samps)

blood_plot <- filter(across_tissue_out, case_when(
    SUBJID %in% top_subj | SUBJID %in% bot_subj ~ TRUE,
    TRUE ~ FALSE
)) %>% mutate(
    top_o_bot = case_when(
        SUBJID %in% top_subj ~ "top",
        SUBJID %in% bot_subj ~ "bot"
    )
)

summ_blood_plot <- blood_plot %>%
    group_by(tissue, top_o_bot) %>%
    summarize(zscores = mean(zscores), top_o_bot) %>%
    unique() 

overall_comparison <- summ_blood_plot %>%
    pivot_wider(id_cols = tissue, names_from = top_o_bot, values_from = zscores) %>%
    mutate(top_over_bot = top > bot,
           mean_diff = (top-bot)) %>%
    filter(!is.na(mean_diff)) %>%
    arrange(desc(mean_diff))


## Plot changes
## Remove where there are no top and bots
## order on mean difference
blood_plot <- blood_plot %>%
    filter(tissue %in% overall_comparison$tissue) %>%
    mutate(tissue = factor(tissue, levels = overall_comparison$tissue))
summ_blood_plot <- summ_blood_plot %>%
    filter(tissue %in% overall_comparison$tissue) %>%
    mutate(tissue = factor(tissue, levels = overall_comparison$tissue))

ggplot(blood_plot, aes(tissue, zscores, color = top_o_bot))  +
    geom_point(aes(group = SUBJID), alpha = .5) +
    geom_line(aes(group = SUBJID), alpha = .5) +
    theme_classic() + 
    theme(axis.text.x = element_text(angle = 80, hjust=1)) +
    geom_boxplot(data = summ_blood_plot, aes(tissue, zscores))

print("whole blood b.2 cluster comparison across tissues")
table(overall_comparison$top_over_bot)
```
How often do these genes show up with HBA1?
```{r}
hba1_overlap <- ClusterGeneOverlaps(c("HBA1", b2_genes),all_clusters = list_dat,only_hits = T) %>%
    mutate(hba_plus = ifelse(str_detect(found_genes, "HBA1") &
                                 str_detect(found_genes, ";"), TRUE,
                             FALSE))
table(hba1_overlap$hba_plus)
```
Do the genes correlate with HBA1 via subclusters?
```{r}

results_out <- lapply(b2_genes, DoGeneSubclustersCorrelate, gene_2 = "HBA1", cluster_list = list_dat, profile_loc_or_frame= "../../MARCC output/variance_profiles_kendall/")

analysis_out <- lapply(results_out, function(x){
    n_of_na <- sum(is.na(x$correlation))
    n_pos_cor <- sum(x$correlation > 0,na.rm = T)
    n_neg_cor <- sum(x$correlation < 0,na.rm = T)
    n_same_clust <- sum(x$correlation == 1,na.rm = T)
    n_total_tissue <- nrow(x)
    res <- cbind(n_same_clust, n_pos_cor, n_neg_cor, n_of_na, n_total_tissue)
    rownames(res) <- colnames(x)[3]
    return(res)
}) 

final_frame <- NULL
for (cur_ind in seq_along(analysis_out)) {
    final_frame <- rbind(final_frame, analysis_out[[cur_ind]])
}
final_frame <- final_frame %>%
    as.data.frame() %>%
    rownames_to_column(var = "gene_name") %>%
    mutate(normal_cor = ifelse(n_pos_cor > n_neg_cor, "positive", "negative"))

final_frame_blood <- final_frame

table(final_frame_blood$normal_cor)
sum(final_frame_blood$n_same_clust)
```

So it appears, overall that genes in B.2 correlate, at least somewhat, wih HBA1. However we know HBA1 correaltes with ischemic time, so how do we know if this is causal or just a correlation? Our next goal is to look at other inflammation related clusters in other tissues and see how often they correlate with HBA1 and across the body. We'll be using the Skin sun exposed cluster C.1 and are going to remove genes that are shared.


Get skin data ready
```{r}
skin_inflammation_genes <- list_dat$skin_sun_exposed__lower_leg$C.1[!(list_dat$skin_sun_exposed__lower_leg$C.1 %in% b2_genes)]


skin_overlaps <- ClusterGeneOverlaps(skin_inflammation_genes, list_dat) %>%
        arrange(desc(n_intersect))

skin_gene_counts <- sort(table(unlist(str_split(skin_overlaps$found_genes,pattern =  "; "))),decreasing = T)

skin_top_genes <- names(skin_gene_counts)[skin_gene_counts > 10]

across_tissue_skin <- GenesFromTissues(genes_of_int = skin_top_genes,
                                      key_dat = key_dat,
                 vst_location = "../../Setup Code/output/",
                 run_all_available = T)

across_tissue_skin$zscores <- NewZScoreMaker(across_tissue_skin)

skin_gene_counts
```


Plot skin data
```{r}
ordered_skin <- across_tissue_skin %>%
    filter(tissue == "skin_sun_exposed__lower_leg") %>%
    arrange(desc(zscores))
top_subj <- head(ordered_skin$SUBJID, n_of_samps)
bot_subj <- tail(ordered_skin$SUBJID, n_of_samps)

skin_plot <- filter(across_tissue_skin, case_when(
    SUBJID %in% top_subj | SUBJID %in% bot_subj ~ TRUE,
    TRUE ~ FALSE
)) %>% mutate(
    top_o_bot = case_when(
        SUBJID %in% top_subj ~ "top",
        SUBJID %in% bot_subj ~ "bot"
    )
) 


summ_skin_plot <- skin_plot %>%
    group_by(tissue, top_o_bot) %>%
    summarize(zscores = mean(zscores), top_o_bot) %>%
    unique() 

overall_comparison <- summ_skin_plot %>%
    pivot_wider(id_cols = tissue, names_from = top_o_bot, values_from = zscores) %>%
    mutate(top_over_bot = top > bot,
           mean_diff = (top-bot)) %>%
    filter(!is.na(mean_diff)) %>%
    arrange(desc(mean_diff))


## Plot changes
## Remove where there are no top and bots
## order on mean difference
skin_plot <- skin_plot %>%
    filter(tissue %in% overall_comparison$tissue) %>%
    mutate(tissue = factor(tissue, levels = overall_comparison$tissue))
summ_skin_plot <- summ_skin_plot %>%
    filter(tissue %in% overall_comparison$tissue) %>%
    mutate(tissue = factor(tissue, levels = overall_comparison$tissue))

ggplot(skin_plot, aes(tissue, zscores, color = top_o_bot))  +
    geom_point(aes(group = SUBJID), alpha = .5) +
    geom_line(aes(group = SUBJID), alpha = .5) +
    theme_classic() + 
    theme(axis.text.x = element_text(angle = 80, hjust=1)) +
    geom_boxplot(data = summ_skin_plot, aes(tissue, zscores))
print("skin inflammation comparison across tissues")
table(overall_comparison$top_over_bot)
```

Do the skin genes correlate with HBA1 via subclusters?
```{r}

results_out <- lapply(skin_top_genes, DoGeneSubclustersCorrelate, gene_2 = "HBA1", cluster_list = list_dat, profile_loc_or_frame= "../../MARCC output/variance_profiles_kendall/")

analysis_out <- lapply(results_out, function(x){
    n_of_na <- sum(is.na(x$correlation))
    n_pos_cor <- sum(x$correlation > 0,na.rm = T)
    n_neg_cor <- sum(x$correlation < 0,na.rm = T)
    n_same_clust <- sum(x$correlation == 1,na.rm = T)
    n_total_tissue <- nrow(x)
    res <- cbind(n_same_clust, n_pos_cor, n_neg_cor, n_of_na, n_total_tissue)
    rownames(res) <- colnames(x)[3]
    return(res)
}) 

final_frame <- NULL
for (cur_ind in seq_along(analysis_out)) {
    final_frame <- rbind(final_frame, analysis_out[[cur_ind]])
}
final_frame <- final_frame %>%
    as.data.frame() %>%
    rownames_to_column(var = "gene_name") %>%
    mutate(normal_cor = ifelse(n_pos_cor > n_neg_cor, "positive", "negative"))

final_frame_skin <- final_frame

table(final_frame_skin$normal_cor)
sum(final_frame_skin$n_same_clust)
```

It appears that the other ischemic related skin cluster (including PLAG2AB) does
not share the same trend with blood perfused tissues. The next goal will be testing
if HBA1 levels as a covariate might help explain the ischemia in the z-score clusters.

