---
title: "Ischemic Time compared to clusters"
author: "Tim Nieuwenhuis"
date: "6/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


```{r}

## Load functions
source("../../global_in/general_scripts.R")

## Master table for tissues
master_key_table <- read.csv("../../automated_cluster_analysis/data_out/key_table.csv")

## get sample data and subject data
subj_dat <- read.csv("../../global_in/gtex_phenotypes_v8.csv")
samp_dat <- read.table(file = "../../global_in/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt",
                       sep = "\t", fill = TRUE, header = TRUE)

```


```{r}
### For each cluster find what percentage of the genes fall into what cell types


### Load in the data
## For each gtex tissue
all_tissues <- unique(master_key_table$r_names)

tissue_dat_list <- list()
for (tiss_ind in seq_along(all_tissues)) {
    tiss_list_out <- list()
    cur_tiss <- all_tissues[tiss_ind]
    tiss_df <- filter(master_key_table, r_names %in% cur_tiss)

    gene_dat <- read.csv(paste0("../../MARCC output/",
                     "variance_genes_kendall/kendall-",
                     cur_tiss,"-gene-clusters-high-variance.csv"))
    
    zscore_dat <- read.csv(paste0("../../MARCC output/",
                                      "variance_profiles_kendall/kendall-",
                                      cur_tiss,"-cluster-profiles.csv"))
    
    tiss_list_out[["gene_dat"]] <- gene_dat
    tiss_list_out[["zscore_dat"]] <- zscore_dat
    tissue_dat_list[[cur_tiss]] <- tiss_list_out
    
}


```

Go through each tissue and do analysis
```{r}

ischemic_sample_dat <- samp_dat %>% select(SAMPID, SMTSISCH)
ischemic_subj_dat <- subj_dat %>% select(SUBJID, TRISCHD)
## For each tissue
cor_table <- NULL
for (tiss_ind in seq_along(tissue_dat_list)) {
    tissue <- names(tissue_dat_list)[[tiss_ind]]
    zscore_dat <- tissue_dat_list[[tissue]][["zscore_dat"]]
    colnames(zscore_dat)[1] <- "SAMPID"
    zscore_dat <- zscore_dat %>%
        mutate(SUBJID = GTEx_SAMPID_to_SUBJID(SAMPID))
    joined_dat <- left_join(zscore_dat, ischemic_subj_dat)
    numeric_ind <- sapply(joined_dat, function(x){is.numeric(x)})
    trisch_ind <- colnames(joined_dat)  == "TRISCHD"
    numeric_dat <- joined_dat[,numeric_ind & !trisch_ind]
    for ( cur_col_ind in seq(ncol(numeric_dat))) {
        cluster <- colnames(numeric_dat)[cur_col_ind]
        new_row <- broom::tidy(cor.test(joined_dat$TRISCHD, numeric_dat[,cur_col_ind]))
        new_row <- cbind(tissue, cluster, new_row)
        cor_table <- rbind(cor_table, new_row)
    }
}

cor_table$p.adj <- p.adjust(cor_table$p.value,method = "holm")

cor_table <- arrange(cor_table, (p.adj))

sum(cor_table$p.adj < 0.05) /nrow(cor_table)
write.csv(cor_table, file = "data_out/all_ischemic_correlations.csv",row.names = F)
```

```{r}
cor_sig <- filter(cor_table, p.adj < 0.05) %>%
    mutate(cluster_dash = str_replace(cluster, "[.]", "-"))

isch_genes <- c()
for (sig_cor in seq(nrow(cor_sig))) {
    cur_tiss <- cor_sig[sig_cor, "tissue"]
    cur_clust <- cor_sig[sig_cor, "cluster"]
    new_genes <- tissue_dat_list[[cur_tiss]][["gene_dat"]][,cur_clust]
    isch_genes <- append(isch_genes, new_genes)
}

isch_table <- as.data.frame(table(isch_genes)) %>% arrange(desc(Freq)) %>%
    filter(isch_genes != "") 
write.csv(isch_table, file = "data_out/top_ischemic_genes.csv",row.names = F)
```


Subset the pseudo correlation on common sig genes and se if they cluster together
or separately 
```{r}
pseudo_cor <- read.csv(file = "../../variation_correlation_pipeline/between_cluster_correlations/data_out/pseudo_correlation_matrix.csv",row.names = 1)
```

```{r}

thresh <- .95
## Limit to top 95% genes
table_95 <- isch_table %>% filter(Freq > quantile(isch_table$Freq, thresh))

ind_95 <- rownames(pseudo_cor) %in% table_95$isch_genes
pseudo_95 <- pseudo_cor[ind_95,ind_95]
pseudo_log_95 <- log2(pseudo_95 + 1)
pdf(file = paste0("data_out/quantile_", thresh,"_ischemic_genes.pdf"), width = 7, height = 5)
pheatmap::pheatmap(pseudo_95,fontsize = 3,border_color = NA)
dev.off()

pdf(file = paste0("data_out/quantile_", thresh,"_ischemic_genes_log2.pdf"), width = 7, height = 5)
pheatmap::pheatmap(pseudo_log_95,fontsize = 4,border_color = NA)
dev.off()
```
Turn pseudo-log into pseudo-proportion
```{r}

pseudo_prop <- pseudo_95

rownames(pseudo_prop) <- str_replace(rownames(pseudo_prop), "-", ".")

horizontal_vals <- apply(pseudo_95, MARGIN = 2, max)

for (row_ind in seq(nrow(pseudo_prop))) {
    row_gene <- rownames(pseudo_prop)[row_ind]
    for (col_ind in seq(ncol(pseudo_prop))) {
       col_gene <- colnames(pseudo_prop)[col_ind] 
       if (horizontal_vals[row_gene] >= horizontal_vals[col_gene]) {
           min_val <- horizontal_vals[col_gene]
       } else{
           min_val <- horizontal_vals[row_gene]
       }
       pseudo_prop[row_ind, col_ind] <- pseudo_prop[row_ind, col_ind]/min_val
        
    }
    
}



pdf(file = paste0("data_out/quantile_", thresh,"_ischemic_genes_proportion.pdf"), width = 7, height = 5)
pheatmap::pheatmap(pseudo_prop,fontsize = 3,border_color = NA)
dev.off()

```




Testing getgene length and gc content
```{r}

gtab <- read.csv("../../global_in/gtab.csv") %>%
    mutate(gene_id_clean = gsub("[.].*",replacement = "",gene_id))

current_genes <- filter(gtab, gene_name %in% row.names(pseudo_95)) %>%
    mutate(gene_id_clean = gsub("[.].*",replacement = "",gene_id))

library(EDASeq)
gclength <- getGeneLengthAndGCContent(current_genes$gene_id_clean,org = "hsa") %>%
    as.data.frame() %>%
    rownames_to_column("gene_id_clean")

gtab_gclength <- left_join(current_genes, gclength)


hist(gtab_gclength$length)
hist(gtab_gclength$gc)
```

## Goal find tissues that have >2 cluster types

Cluster markers
```{r}
blood <- c("HBB")
fosb_mark <- c("FOSB")
adam <- "ADAMTS4"
saa <- "SAA1"

all_genes <- tissue_dat_list$adipose_subcutaneous$gene_dat %>% gather() %>%
    filter(key != "X") %>%
    mutate(in_group = case_when(
        value == blood ~ "blood",
        value == fosb_mark ~ "fosb",
        value == adam ~ "adam",
        value == saa ~ "SAA1",
        TRUE ~ "No clust"
    )) 
pre_clusts <- all_genes %>% select(key, in_group) %>%
    unique()

remove_dupe_ind <- !((duplicated(pre_clusts$key, fromLast = TRUE) | duplicated(pre_clusts$key)) & pre_clusts$in_group == "No clust")

clusts_labeled <- pre_clusts[remove_dupe_ind,] %>% left_join(x = select(all_genes, key, value), y = ., by = "key") %>%
    filter(value != "") %>%
    rename(value = "gene_name")

prepped_genes <- left_join(clusts_labeled, gtab)

gclength_genes <- getGeneLengthAndGCContent(prepped_genes$gene_id_clean,org = "hsa") %>%
    as.data.frame() %>%
    rownames_to_column("gene_id_clean") %>%
    left_join(prepped_genes, y=.)

```

Plot GC content and length distributions
```{r}
ggplot(gclength_genes, aes(gc, color = in_group)) +
    geom_histogram()


ggplot(gclength_genes, aes(length, color = in_group)) +
    geom_histogram()
```


```{r}
pre_plot <- NULL

for (tiss_ind in seq_along(tissue_dat_list)) {
    all_genes <- tissue_dat_list[[tiss_ind]][["gene_dat"]] %>% gather() %>%
        filter(key != "X") %>%
        mutate(in_group = case_when(
            value == blood ~ "blood",
            value == fosb_mark ~ "fosb",
            value == adam ~ "adam",
            value == saa ~ "SAA1",
            TRUE ~ "No clust"
        )) 
    pre_clusts <- all_genes %>% select(key, in_group) %>%
        unique()
    
    remove_dupe_ind <- !((duplicated(pre_clusts$key, fromLast = TRUE) | duplicated(pre_clusts$key)) & pre_clusts$in_group == "No clust")
    
    clusts_labeled <- pre_clusts[remove_dupe_ind,] %>% left_join(x = select(all_genes, key, value), y = ., by = "key") %>%
        filter(value != "") %>%
        rename(value = "gene_name")
    
    prepped_genes <- left_join(clusts_labeled, gtab)
    prepped_genes$tissue <- names(tissue_dat_list)[tiss_ind]
    pre_plot <- rbind(pre_plot, prepped_genes)
    
}

uni_ens <- unique(pre_plot$gene_id_clean)

gclength_genes <- getGeneLengthAndGCContent(uni_ens,org = "hsa") %>%
    as.data.frame() %>%
    rownames_to_column("gene_id_clean") %>%
    left_join(pre_plot, y=.)
```
```{r}
ggplot(gclength_genes, aes(gc, color = in_group)) +
    geom_density()

ggplot(gclength_genes, aes(log2(length), color = in_group)) +
    geom_density()
```

It doesn't appear that ADAMST4, Blood, FOSB, or SAA1 clusters are associated
with GC and gene length associated degredation during ischemic time
