---
title: "Comparing GTEx v8 to GTEx V7"
author: "Tim Nieuwenhuis"
date: "3/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DESeq2)
library(tidyverse)
library(ggforce)
library(viridis)
library(RColorBrewer)
```

```{r}
load("~/Halushka Lab/Matrisome/Matrisome_GTEx_analysis_final/Setup Code/output/gtex-gene-counts-v8.rda")
pheno_dat <- read.csv("~/Halushka Lab/Matrisome/Matrisome_GTEx_analysis_final/global_in/gtex_phenotypes_v8.csv")

short_gtex_vect2vect <- function(x){
  #If else statement used to determine if the name is long or short, if already short nothing should happen, if long, then it will be appropriately shortened. This code works off the assumption that the gtex IDs are either 4 or 5 characters in length

    
    #Generating stand in list
    shortname <- vector()
    #Used subset to create a vector, drop needed to be true to make the tibble output a vector
    for (k in x) {
      #If else for 5 or 4 length IDs
      #Using the "-" to determine if the id has 4 or 5 characters
      if(substr(k, start = 10, stop = 10) == "-"){
        #4 length
        m <- substr(k, 1, 9)
        shortname <- c(shortname, m)
      } else{
        #5 length
        m <- substr(k, 1, 10)
        
        shortname <- c(shortname, m)
        
      }
      
    }
 
  return(shortname)
}
```

Starter comparison between adipose samples
```{r}
res_v7 <- read.csv("~/Halushka Lab/GTEx_variation/variation_tools/gtex_variation_toolbox/v7_output_data/adipose_subcutaneous-gene-clusters-high-variance.csv")

res_v8 <- read.csv("~/Halushka Lab/GTEx_variation/variation_tools/gtex_variation_toolbox/v8_output_data/adipose_subcutaneous-gene-clusters-high-variance.csv")

tissue = "adipose_subcutaneous"
## Need to collapse sub clusters into a single cluster
i <- 1
new_res_v8 <- NULL
while (i <= ncol(res_v8)) {
    ## When there is a column with a [.] combine it with the following column
    sub_clust_bool <- str_detect(colnames(res_v8)[i],pattern = "[.]")
    if (sub_clust_bool == FALSE) {
        new_res_v8 <- as.data.frame(cbind(new_res_v8, res_v8[,i]))
        colnames(new_res_v8)[ncol(new_res_v8)] <- str_sub(colnames(res_v8)[i],start = 1, end = 1)
        i <- i + 1
         next
    }
    joined_col <- c(res_v8[,i], res_v8[,i + 1])
    joined_col <- joined_col[joined_col != ""]
    
    row_diff <- nrow(new_res_v8)  - length(joined_col)
    if (row_diff == 0) {
        new_res_v8 <- as.data.frame(cbind(new_res_v8, joined_col))
        colnames(new_res_v8)[ncol(new_res_v8)] <- str_sub(colnames(res_v8)[i],start = 1, end = 1)
        i <- i + 2
    } else if (row_diff > 0){
        joined_col <- append(joined_col, rep("", row_diff))
        new_res_v8 <- as.data.frame(cbind(new_res_v8, joined_col))
        colnames(new_res_v8)[ncol(new_res_v8)] <- str_sub(colnames(res_v8)[i],start = 1, end = 1)
        i <- i + 2
    } else {
        abs_diff <- abs(row_diff)
        empt_matrix <- matrix(data = "", nrow =abs_diff, ncol = ncol(new_res_v8))
        colnames(empt_matrix) <- colnames(new_res_v8)
        new_res_v8 <- as.data.frame(rbind(new_res_v8, empt_matrix))
        new_res_v8 <- as.data.frame(cbind(new_res_v8, joined_col))
        colnames(new_res_v8)[ncol(new_res_v8)] <- str_sub(colnames(res_v8)[i],start = 1, end = 1)
        i <- i + 2
    }
    ## Use ind + 1 to skip to the next column?
}
res_v7
new_res_v8
```

Understand the overlap:
We need to see what columns have the most overlap and then order them or at least connect them somehow or long format

Long format:
Col 1 is tissue
Col 2 is cluster letter
Col 3 is how many genes it has
Col 5 is how many genes are shared
Col 6 prominent old cluster that shares those genes
Col 7 is how many genes in that specific old cluster are shared


```{r}
comparison_table <- NULL
for (col in seq(ncol(new_res_v8))) {
  vect_col <- new_res_v8[,col]
  new_cluster <- colnames(new_res_v8)[col]
  n_genes_clust <- length(vect_col)
  vect_col <- vect_col[vect_col != ""]
  n_genes_clust <- length(vect_col)
  all_old_clusters <- colSums(
      apply(res_v7,
            MARGIN = 2,
            FUN = function(x){vect_col %in% x}))
  top_old_clust <- all_old_clusters[order(all_old_clusters, decreasing = T)][1]
  all_overlap_genes <- sum(all_old_clusters)
  new_row <- c(tissue,
                new_cluster,
                n_genes_clust,
                all_overlap_genes,
                names(top_old_clust),
                unname(top_old_clust)
                )
  comparison_table <- rbind(comparison_table, new_row)
  }

```

Run all tissue analysis
```{r}
tiss_list <- read.csv("~/Halushka Lab/Matrisome/Matrisome_GTEx_analysis_final/GTEx_age_sex_analysis/input/general_list_test_clean.txt")
comparison_table_v8 <- NULL
comparison_table_v7 <- NULL
for (tissue in tiss_list[,1]) {
    if (startsWith(tissue, "cells_cultured_fi")|
        startsWith(tissue, "colon_tr")|
        startsWith(tissue, "test")) {
        next
    }
    new_path <- paste0("~/Halushka Lab/GTEx_variation/variation_tools/gtex_variation_toolbox/v8_output_data/",
                       tissue,
                       "-gene-clusters-high-variance.csv")
    old_path <- paste0("~/Halushka Lab/GTEx_variation/variation_tools/gtex_variation_toolbox/v7_output_data/",
                       tissue,
                       "-gene-clusters-high-variance.csv")
    
    ##First loop to fix v8 to be like v7
    res_v8 <- read.csv(new_path)
    res_v7 <- read.csv(old_path)
    
    i <- 1
    new_res_v8 <- NULL
    while (i <= ncol(res_v8)) {
        ## When there is a column with a [.] combine it with the following column
        sub_clust_bool <- str_detect(colnames(res_v8)[i],pattern = "[.]")
        if (sub_clust_bool == FALSE) {
            new_res_v8 <- as.data.frame(cbind(new_res_v8, res_v8[,i]))
            colnames(new_res_v8)[ncol(new_res_v8)] <- str_sub(colnames(res_v8)[i],start = 1, end = 1)
            i <- i + 1
             next
        }
        joined_col <- c(res_v8[,i], res_v8[,i + 1])
        joined_col <- joined_col[joined_col != ""]
        
        row_diff <- nrow(new_res_v8)  - length(joined_col)
        if (row_diff == 0) {
            new_res_v8 <- as.data.frame(cbind(new_res_v8, joined_col))
            colnames(new_res_v8)[ncol(new_res_v8)] <- str_sub(colnames(res_v8)[i],start = 1, end = 1)
            i <- i + 2
        } else if (row_diff > 0){
            joined_col <- append(joined_col, rep("", row_diff))
            new_res_v8 <- as.data.frame(cbind(new_res_v8, joined_col))
            colnames(new_res_v8)[ncol(new_res_v8)] <- str_sub(colnames(res_v8)[i],start = 1, end = 1)
            i <- i + 2
        } else {
            abs_diff <- abs(row_diff)
            empt_matrix <- matrix(data = "", nrow =abs_diff, ncol = ncol(new_res_v8))
            colnames(empt_matrix) <- colnames(new_res_v8)
            new_res_v8 <- as.data.frame(rbind(new_res_v8, empt_matrix))
            new_res_v8 <- as.data.frame(cbind(new_res_v8, joined_col))
            colnames(new_res_v8)[ncol(new_res_v8)] <- str_sub(colnames(res_v8)[i],start = 1, end = 1)
            i <- i + 2
        }
        ## Use ind + 1 to skip to the next column?
    }
    
    
        for (col in seq(ncol(res_v7))) {
        vect_col <- res_v7[,col]
        new_cluster <- colnames(res_v7)[col]
        n_genes_clust <- length(vect_col)
        vect_col <- vect_col[vect_col != ""]
        n_genes_clust <- length(vect_col)
        all_old_clusters <- colSums(
          apply(new_res_v8,
                MARGIN = 2,
                FUN = function(x){vect_col %in% x}))
        top_old_clust <- all_old_clusters[order(all_old_clusters, decreasing = T)][1]
        all_overlap_genes <- sum(all_old_clusters)
        new_row <- c(tissue,
                    new_cluster,
                    n_genes_clust,
                    all_overlap_genes,
                    names(top_old_clust),
                    unname(top_old_clust)
                    )
        comparison_table_v7 <- rbind(comparison_table_v7, new_row)
        }
    
        for (col in seq(ncol(new_res_v8))) {
        vect_col <- new_res_v8[,col]
        new_cluster <- colnames(new_res_v8)[col]
        n_genes_clust <- length(vect_col)
        vect_col <- vect_col[vect_col != ""]
        n_genes_clust <- length(vect_col)
        all_old_clusters <- colSums(
          apply(res_v7,
                MARGIN = 2,
                FUN = function(x){vect_col %in% x}))
        top_old_clust <- all_old_clusters[order(all_old_clusters, decreasing = T)][1]
        all_overlap_genes <- sum(all_old_clusters)
        new_row <- c(tissue,
                    new_cluster,
                    n_genes_clust,
                    all_overlap_genes,
                    names(top_old_clust),
                    unname(top_old_clust)
                    )
        comparison_table_v8 <- rbind(comparison_table_v8, new_row)
        }
    
    }

comparison_table <- as_tibble(comparison_table)

# Long format:
# Col 1 is tissue
# Col 2 is cluster letter
# Col 3 is how many genes it has
# Col 5 is how many genes are shared
# Col 6 prominent old cluster that shares those genes
# Col 7 is how many genes in that specific old cluster are shared


comparison_table_v8 <- comparison_table_v8 %>% as_tibble() %>% mutate(V3 = as.numeric(V3),
                           V4 = as.numeric(V4),
                           V6 = as.numeric(V6))

comparison_table_v7 <- comparison_table_v7 %>% as_tibble() %>% mutate(V3 = as.numeric(V3),
                           V4 = as.numeric(V4),
                           V6 = as.numeric(V6))

colnames(comparison_table_v8) <- c("tissue",
                                "v8_cluster",
                                "n_v8_genes",
                                "shared_v8_v7",
                                "match_v7_clust",
                                "match_v7_genes"
                                )

colnames(comparison_table_v7) <- c("tissue",
                                "v7_cluster",
                                "n_v7_genes",
                                "shared_v7_v8",
                                "match_v8_clust",
                                "match_v8_genes"
                                )




```


Plots: version_difference clusters per tissue 
Plots: version_difference genes per tissue
```{r}
## Genes in v7 or v8 and not in the other
comparison_table_v8 <- comparison_table_v8 %>% mutate(missing_in_v7 = n_v8_genes - shared_v8_v7,
                                                      analysis = "v8")
comparison_table_v7 <- comparison_table_v7 %>% mutate(missing_in_v8 = n_v7_genes - shared_v7_v8,
                                                      analysis = "v7")

summaried_v8 <- comparison_table_v8 %>%
    group_by(tissue) %>%
    summarise(unique_genes = sum(missing_in_v7),
              n_clust = n()) %>%
    mutate(analysis = "v8")

summaried_v8_nx <- comparison_table_v8 %>%
    group_by(tissue) %>%
    filter(v8_cluster != "X") %>%
    summarise(unique_genes = sum(missing_in_v7),
              n_clust = n()) %>%
    mutate(analysis = "v8")


summaried_v7 <- comparison_table_v7 %>%
    group_by(tissue) %>%
    summarise(unique_genes = sum(missing_in_v8),
              n_clust = n()) %>%
    mutate(analysis = "v7")

summaried_v7_nx <- comparison_table_v7 %>%
    group_by(tissue) %>%
    filter(v7_cluster != "X") %>%
    summarise(unique_genes = sum(missing_in_v8),
              n_clust = n()) %>%
    mutate(analysis = "v7")

summaried_dat <- rbind(summaried_v8,summaried_v7)
summaried_dat_nx <- rbind(summaried_v8_nx,summaried_v7_nx)


ggplot(summaried_dat, aes(y= tissue, x = n_clust, fill = analysis)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_viridis(discrete=TRUE) +
    labs(title = "Number of clusters per tissue across analyses") +
    xlab("Number of clusters") +
    ylab("Tissues") +
    theme_classic()
ggsave(filename = "output/cluster_n_version_comparison.pdf", height = 7.5)


ggplot(summaried_dat, aes(y= tissue, x = unique_genes, fill = analysis)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_viridis(discrete=TRUE) +
    labs(title = "N of genes not shared between analyses") +
    xlab("Genes Unique to Analysis") +
    ylab("Tissues") +
    theme_classic()
ggsave(filename = "output/gene_unique_to_analysis.pdf", height = 7.5)


ggplot(summaried_dat_nx, aes(y= tissue, x = unique_genes, fill = analysis)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_viridis(discrete=TRUE) +
   # scale_color_discrete( brewer.pal(7, "BrBG"), guide = FALSE) +
    labs(title = "N of genes not shared between analyses",
         subtitle = "No X cluster") +
    xlab("Genes Unique to Analysis") +
    ylab("Tissues") +
    theme_classic()
ggsave(filename = "output/gene_unique_to_analysis_no_x.pdf", height = 7.5)


```

```{r}
sink(file = paste0('run_history/cluster_comparison_session_info_', Sys.Date(), ".txt"),
     type = "output")
sessionInfo()
sink()
```

