---
title: "Post-curate Recreator"
author: "Tim N"
date: "12/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this markdown is multifaceted, but the overall point is to update
tissues where we have manually collapsed their clusters. Therefore there are 3
sections of code required.

1. Rename the clusters (just so its in alphabetical order again)
2. Recreate Z-scores
3. Recreate gene list
4. Make new key dat

Then I need to re-run all my analyses that use Z-scores, but that won't be
done through this code.

Load data and code I think will be required
```{r}
key_dat <- read.csv("../automated_cluster_analysis/data_out/key_table.csv")
source("../global_in/general_scripts.R")
source("../global_in/pipeline_specific_functions/mccall_analysis_corr_step.R")
source("../global_in/pipeline_specific_functions/mccall_cluster_generator.R")
source("../global_in/pipeline_specific_functions/sub_cluster_maker.R")
source("../global_in/pipeline_specific_functions/mccall_profile_maker.R")
source("../global_in/pipeline_specific_functions/unfiltered_heatmap.R")
correlation_type = "kendall"
library(tidyverse)
require(DESeq2)
library(pheatmap)
library(RColorBrewer)
library(SuppDists)
library(reshape2)
```



Load in curated files and select what tissues will be re-analyzed
```{r}
old_manual_dat <- readxl::read_xlsx("../manual_cluster_analysis/manual_curation_12_17_21_ton.xlsx",
                                    sheet = 2) %>%
    separate(`GTEx label`, c("abbreviation", "cluster", "sub_num"), "[.]",remove = F)
curated_dat <- filter(old_manual_dat, curated == 1)

key_dat_cur <- select(key_dat, abrreviated, r_names, official_name) %>%
    filter(abrreviated %in% curated_dat$abbreviation) %>%
    unique()
```

Make a list that will store the changes for each tissue so we can do this via
a for loop.

```{r}
curation_list <- list()
```
Fix letters
```{r}
new_manual_dat <- filter(old_manual_dat, !(abbreviation %in% key_dat_cur$abrreviated))
to_rejoin <- NULL
for (curated_ind in seq(nrow(key_dat_cur))) {
    curated_row <- key_dat_cur[curated_ind,]
    curated_rname <- curated_row$r_names
    curated_abbreviation <- curated_row$abrreviated
    current_dat <- old_manual_dat %>% filter(abbreviation %in% curated_abbreviation)
    ## for loop renaming clusters
    alpha_ind <- 1
    cluster_vect <- current_dat$cluster
    pair <- FALSE
    new_cluster_vect <- current_dat$cluster
    for (cluster_ind in seq((cluster_vect))) {
        if (pair == TRUE | is.na(pair)) {
            pair = FALSE
            next()
        }
        ## Catch sub clusters
        cur_letter <- cluster_vect[cluster_ind]
        next_letter <- cluster_vect[cluster_ind + 1]
        pair <- cur_letter == next_letter
        pair <- if_else(is.na(pair), FALSE, pair)
        ## Change lettering if needed
        if (pair == TRUE) {
            new_cluster_vect[cluster_ind:(cluster_ind + 1)] <- LETTERS[alpha_ind]
            alpha_ind = alpha_ind + 1
        } else{
            new_cluster_vect[cluster_ind] <- LETTERS[alpha_ind]
            alpha_ind = alpha_ind + 1
        }
        
        
    }
    current_dat$cluster <- new_cluster_vect
    to_rejoin <- rbind(to_rejoin, current_dat)
}

new_manual_dat <- rbind(new_manual_dat,to_rejoin) 
```


```{r}

for (cur_key_ind in seq(nrow(key_dat_cur))) {
    cur_row <- key_dat_cur[cur_key_ind,]
    temp_manual_dat <- new_manual_dat %>% filter(abbreviation %in% cur_row$abrreviated)
    ##### Run first half of analysis here
    pipeline_dat_1 <- mccall_analysis_corr_step(cur_row$r_names,
                                     path_to_files = "../Setup Code/output/",
                                     variance_type = "quantile",
                                     variance_value = .98,
                                     correlation = correlation_type,
                                     MARCC = FALSE)
    pipeline_dat_2 <- mccall_cluster_generator(pipeline_dat_1,
                                         cluster_breaker = TRUE,
                                         sub_clust_cut_off = .8,
                                         agglo_cut_off = .75)
    ####
    old_anno <- pipeline_dat$anno_df
    new_anno <- pipeline_dat$anno_df
    for (cluster_ind in seq(nrow(temp_manual_dat))) {
        cur_clust_row <- temp_manual_dat[cluster_ind,]
        cur_genes <- str_split(cur_clust_row$genes, ";")[[1]]
        clust_ind <- rownames(old_anno) %in% cur_genes
        new_anno[clust_ind,1] <- cur_clust_row$cluster
    }
    
    new_anno$cluster <- factor(new_anno$cluster,
                               levels =  c("",
                                           LETTERS[1:length(unique(temp_manual_dat$cluster))]))
    pipeline_dat_2$anno_df <- new_anno
    pipeline_dat_2$cluster_names <- as.character(new_anno$cluster)
    pipeline_dat_3 <- sub_cluster_maker(pipeline_dat_2)
    pipeline_dat_4 <- mccall_profile_maker(pipeline_dat_3)
    unfiltered_heatmap(pipeline_dat_4,
                   save_file = paste0("new_heatmaps/curated_",
                                      cur_row$r_names,
                                      ".pdf"),
                   row_col_font = 1.5) 
    profiles_path <- paste0("new_zscores/",
                        correlation_type, "-",cur_row$r_names,"-cluster-profiles.csv")
gene_clust_path <- paste0("new_gene_lists/",
                          correlation_type, "-",cur_row$r_names,"-gene-clusters-high-variance.csv")
    write.csv(pipeline_dat_4[["sample_profiles"]], file=profiles_path, quote=FALSE)
    write.csv(pipeline_dat_4[["cluster_genes"]], file = gene_clust_path,
          row.names = FALSE)
}

```



```{r}
sub_cluster_maker <- function(clust_initial){
    ### This function splits the clusters created by absolute values into two
    ### seperate cluster eg. X-1 and X-2. This is to make the clusters more
    ### informative. 
    ## Load in the data
    cnames <- clust_initial$cluster_names
    anno <- clust_initial$anno_df
    gcor <- clust_initial$correlation_matrix
    vsdHighVar <- clust_initial[["high_var_genes"]]
    ## Generate gclusts matrix, which includes the genes and their respective
    ## clusters.
    gclusts <- matrix("", nrow=max(table(cnames)), ncol=length(table(cnames)))
    colnames(gclusts) <- levels(anno$cluster)
    for(k in 1:length(table(cnames))){
        tmp <- colnames(gclusts)[k]
        cur_genes <- rownames(anno)[anno$cluster == tmp]
        #gclusts[1:sum(cnames==tmp), k] <- rownames(gcor)[cnames==tmp]
        gclusts[1:sum(cnames==tmp), k] <- cur_genes
        
    }
    
    ## Below code creates subclusters, splitting on the absolute values
    for (column in LETTERS[1:(length(unique(cnames)) -1 )]) {
        #Pull out the first column
        test_col <- gclusts[,column]
        first_gene <- test_col[1]
        #Make an index to subset gcor to just a single clusters
        ind <- which(rownames(gcor) %in% test_col)
        sub_cor <- gcor[ind,ind]
        #Test if the cluster has any negatively correlating genes
        ## This doesn't work let me try a new method generating a new dendo
        ind2 <- sub_cor[,first_gene] < 0
        #If there are negatively correlating genes...
        if (sum(ind2) > 0) {
            dendo_sub <- hclust(d=as.dist(1-(sub_cor)), method = "average")
            
            ## clean up heatmap to focus on interesting gene clusters
            sub_clusts <- cutree(dendo_sub,k = 2)
            
            #generate sub cluster 1
            sub_1 <- names(sub_clusts)[sub_clusts == 1]
            #Subcluster 2
            sub_2 <- names(sub_clusts)[sub_clusts == 2]
            #Remover old column to replace with new columns
            gclusts <- as.matrix(gclusts[,!(colnames(gclusts) == column)])
            #Make a single matrix of same rows as gclusts, one for each sub-cluster. 
            matrix_1  <- as.matrix(c(sub_1 ,rep("", nrow(gclusts) - length(sub_1))), )
            colnames(matrix_1) <- paste0(column,"-1")
            matrix_2  <- as.matrix(c(sub_2 ,rep("", nrow(gclusts) - length(sub_2))), )
            colnames(matrix_2) <- paste0(column,"-2")
            #Add subclusters to gclusts
            gclusts <-   cbind(gclusts, matrix_1, matrix_2)
            
        }
        
    }
    
    #Realphebatize columns for sake of beauty 
    gclusts <- gclusts[,order(colnames(gclusts))]
    
    ## Stratify cluster inclusion
    ind <- which(rownames(vsdHighVar) %in% rownames(anno)[anno$cluster != ""])
    vsdHighVarClust <- clust_initial$high_var_genes[ind,]
    vsdHighVarCenteredClust <- clust_initial$high_var_centered[ind,]
    vsdHighVarNoise <- clust_initial$high_var_genes[-ind,]
    
    ## Return output to the input list
    clust_initial[["cluster_genes"]] <- gclusts
    clust_initial[["high_var_clusters"]] <- vsdHighVarClust
    clust_initial[["high_var_clust_cent"]] <- vsdHighVarCenteredClust
    clust_initial[["high_var_noise"]] <- vsdHighVarNoise
    clust_initial
}
```

