---
title: "Age, sex, ischemic curation inclusion"
author: "Tim Nieuwenhuis"
date: "7/14/2021"
output: html_document
---
#This file is to only be run once after runing all_batch_analysis.rmd!

The point of this file is to edit the document of curated cluster analysis
(manual_curation.csv). We will add two new columns with a boolean saying if
a given cluster is or is not associated with either sex or ischemic time.
This is only to be ran once because automated change to manual_curation.csv could
lead to unknown issues as it is genreally a manually changed file.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

load data
```{r}
curation_dat <- read.csv(file = "../../manual_cluster_analysis/manual_curation.csv")
ischemic_dat <- read.csv(file = "data_out/ischemic_batch_results.csv")
sex_dat <- read.csv(file = "data_out/sex_batch_results.csv")
key_dat <- read.csv("../../automated_cluster_analysis/data_out/key_table.csv")
```

Prepare sex and ischemic clusters to interact with curation dat, by including
abbreviations and joining clusters to abbreviations
```{r}
cur_keys <- key_dat %>% select(label, cluster, tissue=  "r_names")
sex_joined <- left_join(sex_dat, cur_keys, by = c("tissue", "cluster"))
ischemic_joined <- left_join(ischemic_dat, cur_keys, by = c("tissue", "cluster"))
```
Mark curation dat if sig_columns exist in 
```{r}
sig_ischemic <- filter(ischemic_joined, p_adj < 0.05) %>%
    select(label,isch_estimate = estimate)
sig_sex <- filter(sex_joined, p_adj < 0.05)%>%
    select(label,sex_estimate = estimate)

curation_dat <- curation_dat %>% left_join(sig_ischemic) %>% left_join(sig_sex)
curation_dat <- curation_dat %>% mutate(ischemic_sig = case_when(
    isch_estimate > 0 ~ "positive_corr",
    isch_estimate < 0 ~ "negative_corr",
    TRUE ~ "no_corr"),
    sex_sig = case_when(
        sex_estimate > 0 ~ "positive_corr",
        sex_estimate < 0 ~ "negative_corr",
        TRUE ~ "no_corr"
        )
    ) %>% select(-sex_estimate, -isch_estimate)


# curation_dat$ischemic_sig <- curation_dat$label %in% sig_ischemic$label
# curation_dat$sex_sig <- curation_dat$label %in% sig_sex$label

write.csv(curation_dat, file = "../../manual_cluster_analysis/manual_curation.csv",row.names = F)
```

