---
title: "Different tests"
author: "Tim Nieuwenhuis"
date: "4/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Variables:
vst vs variance stabilizing transformation
old gtex data vs new

Load in best data
```{r}
golden_goose <- read.csv("../../../GTEx_variation/v8_output/lung-gene-clusters-high-variance.csv")
load("data_in/lung-vsd-mean-filtered.rda")

```


Initially we clean and prepare the old and new data

old data pipe 
```{r}
load(file = "data_in/gtex-gene-counts-v8-old.rda")
samp_ind <- (stab$SMTSD == "Lung")
dat_lung <- dat[,samp_ind]
stab_lung <- stab[samp_ind,]
old_gtab <- gtab
## Make DESeq object
des_dat_old <- DESeqDataSetFromMatrix(countData = dat_lung,
                                  colData = stab_lung,
                                  design = ~1)
```

New data pipe
```{r}
rm(dat, gtab, stab)
load(file = "../global_in/gtex-gene-counts-v8.rda")
samp_ind <- (stab$SMTSD == "Lung")
dat_lung <- dat[,samp_ind]
stab_lung <- stab[samp_ind,]
new_gtab <- gtab
## Make DESeq object
des_dat_new <- DESeqDataSetFromMatrix(countData = dat_lung,
                                  colData = stab_lung,
                                  design = ~1)
```

Variance stabilizing transformation
```{r} 
full_old_norm <- varianceStabilizingTransformation(des_dat_old)
full_new_norm <- varianceStabilizingTransformation(des_dat_new)

```

vst
```{r}
vst_old_norm <- vst(des_dat_old)
vst_new_norm <- vst(des_dat_new)
```
Put everything into one list for the for loop
```{r}
normalized_list <- list(old_vst = vst_old_norm,
                             new_vst = vst_new_norm,
                             old_full = full_old_norm,
                             new_full = full_new_norm)
```


Loop through th matt mcall analysis to see which one works

```{r}
filtered_normalized <- list()
filtered_gtab <- list()
for (test_ind in seq(length(normalized_list))) {
    analysis_name <- names(normalized_list[test_ind])
    if (str_detect(analysis_name, "new")) {
        gtab <- new_gtab
    } else {
        gtab <- old_gtab
    }
    
    
    dat <- assay(normalized_list[[test_ind]])
    m <- rowMeans(dat)
    ind <- which(m>5)
    gtab_filt <- gtab[ind,]
    dat_filt <- dat[ind,]
    filtered_normalized[[analysis_name]] <- dat_filt
    filtered_gtab[[analysis_name]] <- gtab_filt
}

```
Compare filtered data to the data we load in
```{r}
dim(assay(lungVSDMeanFiltered))

for ( test_ind in seq(length(filtered_normalized))) {
    print(names(filtered_normalized[test_ind]))
    print("Dims")
    res_one <- dim(assay(lungVSDMeanFiltered)) == dim(filtered_normalized[[test_ind]])
    print(res_one)
    print("Total match")
    res_two <- identical((assay(lungVSDMeanFiltered)), filtered_normalized[[test_ind]])
    print(res_two)
}

identical(filtered_normalized$old_full,filtered_normalized$new_full)

filtered_normalized


sum((assay(lungVSDMeanFiltered)) - filtered_normalized$new_full)

assay(lungVSDMeanFiltered)[1:5,1:5]
filtered_normalized$new_full[1:5,1:5]
```

We see that the "full" version is accurate. Is it still accurate after going through the mccall pipeline

```{r}
numeric_var <- 4
for_variance_normalized <- filtered_normalized[3:4]
for_variance_gtab <- filtered_gtab[3:4]
gclusts_list <- list()

for (analysis_ind in seq(length(for_variance_normalized))) {
    analysis_name <- names(for_variance_normalized[analysis_ind])
    vst_lung_mean_filtered <- for_variance_normalized[[analysis_ind]]
    gtab_lung <- for_variance_gtab[[analysis_ind]]
        
        ## Filter down to high variance genes
    v <- apply(vst_lung_mean_filtered,1,var)
    ind <- which(v > numeric_var)
    high_var_lung <- vst_lung_mean_filtered[ind,]
    high_var_gtab <- gtab_lung[ind,]
    high_var_matrix_lung <- (high_var_lung)
    
    # save(vst_lung_mean_filtered, stab_lung, gtab_lung,
    #      high_var_lung,
    #      high_var_gtab,
    #      file = paste0("gene_analysis_comparisons/data_out/lung_vst_var_",cur_var,"-",version,".rda"))
    # save(vst_lung_mean_filtered, stab_lung, gtab_lung,
    #      high_var_lung,
    #      high_var_gtab,
    #      file =  paste0("gene_analysis_comparisons/data_in/lung_vst_var_",cur_var,"-",version,".rda"))
    # 
    ## Alphnum clustering method
    rownames(high_var_matrix_lung) <- high_var_gtab$gene_name
    vsd_high_var_centered_lung <- high_var_matrix_lung - rowMeans(high_var_matrix_lung)
    
    gcor <- cor(t(high_var_matrix_lung))
    
    ## Generate critical value
    ngene <- nrow(high_var_matrix_lung)
    nsamp <- ncol(high_var_matrix_lung)
    tauCrit <- round(qKendall(0.01/((ngene^2)-ngene)/2, nsamp, lower.tail = F), 2)
    
    ## cluster based on absolute correlation distance
    dendo <- hclust(d=as.dist(1-abs(gcor)), method = "average")
    
    ## clean up heatmap to focus on interesting gene clusters
    clusts <- cutree(dendo, h=1-tauCrit)
    tabclusts <- table(clusts)
    ind <- which(tabclusts >= 6) #arbitrary
    clusts[!(clusts %in% names(tabclusts)[ind])] <- 0
    ntmp <- names(clusts)
    clusts <- as.numeric(as.factor(clusts))-1
    names(clusts) <- ntmp
    
    
    #Naming clusters based on their clust values
    #Very manual process
    n <- length(table(clusts))
    cnames <- rep("", length(clusts))
    
    #Tim for loop
    for (clst in 1:(n)){
        cnames[clusts==clst] <- LETTERS[clst]
    }
    
    anno = data.frame("cluster"=factor(cnames))
    rownames(anno) = names(clusts)
    ## make matrix to output
    gclusts <- matrix("", nrow=max(table(cnames)), ncol=length(table(cnames)))
    colnames(gclusts) <- levels(anno$cluster)
    for(k in 1:length(table(cnames))){
        tmp <- colnames(gclusts)[k]
        gclusts[1:sum(cnames==tmp), k] <- rownames(gcor)[cnames==tmp]
    }
    
    #Sub clusters code! Now annotated
    
    for (column in LETTERS[1:(length(unique(cnames)) -1 )]) {
        #Pull out the first column
        test_col <- gclusts[,column]
        first_gene <- test_col[1]
        #Make an index to subset gcor to just a single clusters
        ind <- which(rownames(gcor) %in% test_col)
        sub_cor <- gcor[ind,ind]
        #Test if the cluster has any negatively correlating genes
        ind2 <- sub_cor[,first_gene] < 0
        
        #If there are negatively correlating genes...
        if (sum(ind2) > 0) {
            #generate sub cluster 1
            sub_1 <- rownames(sub_cor)[!ind2]
            #Subcluster 2
            sub_2 <- rownames(sub_cor)[ind2]
            
            #Remover old column to replace with new columns
            gclusts <- as.matrix(gclusts[,!(colnames(gclusts) == column)])
            
            
            #nrow(gclusts) Not sure if useful, keeping for now
            #add sub-columns
            
            #Make a single matrix of same rows as gclusts, one for each sub-cluster. 
            matrix_1  <- as.matrix(c(sub_1 ,rep("", nrow(gclusts) - length(sub_1))), )
            colnames(matrix_1) <- paste0(column,"-1")
            
            # cbind(gclusts, matrix_1, matrix_2)
            
            
            matrix_2  <- as.matrix(c(sub_2 ,rep("", nrow(gclusts) - length(sub_2))), )
            colnames(matrix_2) <- paste0(column,"-2")
            
            #Add subclusters to gclusts
            gclusts <-   cbind(gclusts, matrix_1, matrix_2)
            
        }
        
        
    }
    
    
    
    #Realphebatize columns for sake of beauty 
    gclusts <- gclusts[,order(colnames(gclusts))]
    
    gclusts_list[[analysis_name]] <- gclusts
}



```


```{r}
length(unique(unlist(golden_goose)))
length(unique(unlist(gclusts_list$new_full)))

df_out <- gclusts_list$new_full

"SFTPB" %in% df_out[,"B-1"]


```

