---
title: "cogaps_lung_analysis"
author: "Tim Nieuwenhuis"
date: "3/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(CoGAPS)
library(DESeq2)
library(broom)
library(tidyverse)
```

Select data usomg var options
```{r}
## Select variant
cur_var <- 2.5
```


Load data
```{r}
paste0("data_in/lung_vst_var_",cur_var,".rda")
#load("data_out/lung_vst")
load(paste0("data_in/lung_vst_var_",cur_var,".rda"))
lung_clusts <-  read.csv(paste0("data_out/lung-gene-clusters-",cur_var,"-variance.csv"))
lung_profiles <- read.csv(paste0("data_out/lung-cluster-profiles-",cur_var,"-var.csv"))
n_of_clusts <- ncol(lung_clusts) - 1
```

Load Functions
```{r}

join_cogaps_gene_dat <- function(cogaps_obj, gtab){
    ## Bring together cogaps data an gtab
    gene_patterns <- as.data.frame(getAmplitudeMatrix(cogaps_obj)) %>%
        rownames_to_column(var = "gene_id")
    joined_patterns <- left_join(gene_patterns, gtab) %>%
        select(gene_id, gene_name, everything())
    return(joined_patterns)
}

join_cogaps_samp_dat <- function(cogaps_obj, df_profiles){
    sample_patterns_cogaps <- as.data.frame(CoGAPS::getPatternMatrix(cogaps_obj)) %>%
            rownames_to_column(var = "SAMPID")
    sample_patterns_classic <- as.data.frame(df_profiles) %>% rename(SAMPID = "X")

    joined_samp <- left_join(sample_patterns_cogaps, sample_patterns_classic)
    return(joined_samp)
}

cogaps_quicK_check <- function(dat, col_1, col_2, plot = FALSE, cor_test = "kendall"){
    ## Quickly finds correlation between data
    test <- cor.test(dat[[col_1]], dat[[col_2]], method = cor_test)    
    tidy_res <- tidy(test)
    tidy_res
    if (plot == TRUE) {
    plt <-  ggplot(dat,  aes_string(x= col_1, y = col_2)) +
            geom_point() +
            labs(title = paste0(col_1, " compared to ", col_2),
                 subtitle = paste0("P-value is ", format.pval(tidy_res$p.value))) +
            theme_classic()
    return(plt)
    }
    return(tidy_res)
}


top_cogaps_genes <- function(dat, pattern_n, n_genes = 10, top_o_bot = "top"){
    if (top_o_bot == "top") {
        pattern <- sym(paste0("Pattern_",pattern_n))
        dat <- dat %>%
            arrange(desc({{pattern }} ))
    } else {
    pattern <- sym(paste0("Pattern_",pattern_n))
    dat <- dat %>%
        arrange({{pattern }})
    }
    
    dat <- dat %>% select(gene_id, gene_name, pattern)
    
    if (n_genes == "all") {
        return(dat)
    }
    
    return(head(dat, n_genes))
    
}

matched_prof_pattern_plots <- function(joined_samps,
                                       cluster_gaps_sum_out,
                                       all_comparisons = TRUE,
                                       sig_plots = TRUE){
    plot_list <- list()
    df_out <- NULL
    ## Check last possible cluster
    ## Move down dataframe until all clusters have at least one selected top hit
    alphanum_clusters <- colnames(joined_samps)[!startsWith(colnames(joined_samps), "Pa") &
                                   !startsWith(colnames(joined_samps), "SA")]
    
    ## Make table of results via rbind
    for (cur_row in seq(nrow(cluster_gaps_sum_out))) {
        cur_row_full <- cluster_gaps_sum_out[cur_row,]
        
        new_row <- cogaps_quicK_check(joined_samps, cur_row_full[[1]],
                                      cur_row_full[[2]], plot = FALSE)
        new_row$comparison <- paste0(cur_row_full[[1]], "-vs-", cur_row_full[[2]])
        df_out <- rbind(df_out, new_row)
        
        }
    if (sig_plots == TRUE) {
        filt_df <- df_out %>% mutate(p.value_adj = p.adjust(p.value, method = "bonferroni")) %>%
            filter(p.value_adj < 0.05) %>%
            arrange(p.value_adj)
        for (row_filt in seq(nrow(filt_df))) {
            split_dat <- str_split(filt_df[row_filt,"comparison"],pattern = "-vs-")
            split_dat[[1]][2]
            plt <- cogaps_quicK_check(joined_samps, split_dat[[1]][1], split_dat[[1]][2], plot = T)
            plot_list[[filt_df[[row_filt,"comparison"]]]] <- plt
        }
        full_obj <- list()
        full_obj[["table"]] <- df_out
        full_obj[["plots"]] <- plot_list
        return(full_obj)
    }
    return(df_out)
}


cluster_cogaps_sum <- function(gene_cluster, cogaps_genes){
    clust_only <- gene_cluster[,-1]
    data_out <- NULL
    for (cluster in seq(ncol(clust_only))) {
        genes <- unique(clust_only[,cluster])
        for (patterns in seq(ncol(cogaps_genes)- 2)) {
            cur_genes <- top_cogaps_genes(cogaps_genes,
                                          patterns, n_genes = "all") %>%
                filter(gene_name %in% genes)
            cur_res <- mean(cur_genes[,3])
            data_out <- rbind(data_out,
                c(colnames(clust_only)[cluster], paste0("Pattern_",patterns), cur_res))
        }
    }
    colnames(data_out) <- c("Clusters", "Patterns", "Mean Gene Association From Pattern")
    data_out <- as.data.frame(data_out) %>% arrange(desc(`Mean Gene Association From Pattern`))
    data_out[,3] <- as.numeric(data_out[,3])
    return(data_out)
    }

cogaps_alphnum_comparison_df <- function(cogs_obj,
                                         alphnum_clusts,
                                         alphnum_profiles,
                                         gtab){
    ## Incorporate previous functions to make this require less functions
    
    sample_patterns <- join_cogaps_samp_dat(cogs_obj, lung_profiles)
    gene_patterns <- join_cogaps_gene_dat(cogs_obj, gtab)
    alphnum_cog_matched <- cluster_cogaps_sum(alphnum_clusts, gene_patterns)
    ## New code
    final_out <- NULL
    pval_df <- matched_prof_pattern_plots(sample_patterns, alphnum_cog_matched, sig_plots = F) %>%
        separate(comparison, c("alphnum", "cog_pattern"),sep = "-vs-", remove = FALSE)
    ## For loop for all patterns
    pattern_marks <- patternMarkers(cogs_obj)
    pattern_marks <- lapply(pattern_marks$PatternMarkers, function(x){
                        ind <- match(x, gtab$gene_id)
                        renamed_dat <- gtab$gene_name[ind]
    
                      })
    patterns <- names(pattern_marks)
    for (pattern in patterns) {
        #Col 1 is pattern
        col_1 <- pattern
        #Col 2 get best correlating alphnum
        top_res <- filter(pval_df, cog_pattern == pattern) %>%
            arrange(p.value) %>% slice(1)
        col_2 <- top_res$alphnum
        #Col 3 p value
        col_3 <- top_res$p.value
        # Col 4 pattern from cogaps or ours
        markers <- pattern_marks[[pattern]]
        
        col_4 <- ifelse(is_empty(markers), "cogaps_failed", "cogaps_worked")
        #Col 5 genes from patterns
        if (is_empty(markers)) {
            ### Pull top genes from alphnum data
            cogaps_genes <- join_cogaps_gene_dat(cogs_obj, gtab)
            alphnum_genes <- top_cogaps_genes(cogaps_genes,
                                             pattern_n = as.integer(stringr::str_extract(pattern, "\\d+")),
                                             n_genes = 10)
            ## In the future maybe come up with a filter on the top associated genes
            col_5 <- alphnum_genes$gene_name
             col_5 <- paste0(col_5, collapse = ", ")
        } else {
            col_5 <- pattern_marks[[pattern]]
            col_5 <- paste0(col_5, collapse = ", ")
        }
        ## Col 6 genes from most associatd alphnum clust
        col_6 <- alphnum_clusts[[top_res$alphnum]]
        col_6 <- col_6[col_6 != ""]   
        col_6 <- paste0(col_6, collapse = ", ")
    
            new_row <- as.data.frame(cbind(col_1, col_2, col_3, col_4, col_5, col_6))
            colnames(new_row) <- c("CoGAPS Pattern",
                                   "Alphnum Cluster",
                                   "P-value of sample correlation",
                                   "Can cogaps take out marker genes?",
                                   "Cogaps Pattern Markers",
                                   "Alphnum Cluster Genes")
            final_out <- rbind(final_out, new_row)
    }
    
    return(final_out)
    
}
```



Set parameters
```{r}
params <- new("CogapsParams")
params <- setParam(params, "nIterations", 10000)
params_e <- setParam(params, "nPatterns", n_of_clusts)
params_d <- setParam(params, "nPatterns", n_of_clusts - 3)
params_u <- setParam(params, "nPatterns", n_of_clusts + 3)
```

Matched Clusters
```{r}
cog_out_e <- CoGAPS(assay(high_var_lung),
                  params = params_e)

plot(head(cog_out_e))
```

3 More clusters
```{r}
cog_out_u <- CoGAPS(assay(high_var_lung),
                  params = params_u)

plot(head(cog_out_u))
```

3 Less clusters
```{r}
cog_out_d <- CoGAPS(assay(high_var_lung),
                  params = params_d)

plot(head(cog_out_d))
```

```{r}
cog_out_e
## Gene data
getFeatureLoadings(cog_out_e)[1:5,1:5]
gene_patterns_cogaps_e <- as.data.frame(getAmplitudeMatrix(cog_out_e)) %>%
    rownames_to_column(var = "gene_id")
## Sample data
CoGAPS::getSampleFactors(cog_out_e)[1:5,1:5]
sample_patterns_cogaps_e <- as.data.frame(CoGAPS::getPatternMatrix(cog_out_e)) %>%
            rownames_to_column(var = "SAMPID")
sample_patterns_classic <- as.data.frame(lung_profiles) %>% rename(SAMPID = "X")

joined_samp_e <- left_join(sample_patterns_cogaps_e, sample_patterns_classic)


ggplot(joined_samp_e, aes(A.1, Pattern_7)) +
    geom_point()

joined_gene_e <- left_join(gene_patterns_cogaps_e, gtab_lung)

joined_gene_e %>% filter(gene_name %in% lung_clusts$F.1 | gene_name %in% lung_clusts$F.2) %>%
    ggplot(aes(gene_name, Pattern_7)) +
    geom_point()
    
join_cogaps_gene_dat <- function(cogaps_obj, gtab){
    ## Bring together cogaps data an gtab
    gene_patterns <- as.data.frame(getAmplitudeMatrix(cogaps_obj)) %>%
        rownames_to_column(var = "gene_id")
    joined_patterns <- left_join(gene_patterns, gtab) %>%
        select(gene_id, gene_name, everything())
    return(joined_patterns)
}

join_cogaps_samp_dat <- function(cogaps_obj, df_profiles){
    sample_patterns_cogaps <- as.data.frame(CoGAPS::getPatternMatrix(cogaps_obj)) %>%
            rownames_to_column(var = "SAMPID")
    sample_patterns_classic <- as.data.frame(df_profiles) %>% rename(SAMPID = "X")

    joined_samp <- left_join(sample_patterns_cogaps, sample_patterns_classic)
    return(joined_samp)
}
```
Quick check function
```{r}
dat <- joined_samp_e
col_1<- "Pattern_7"
col_2 <- "F.1"
cogaps_quicK_check <- function(dat, col_1, col_2, plot = FALSE, cor_test = "kendall"){
    ## Quickly finds correlation between data
    test <- cor.test(dat[[col_1]], dat[[col_2]], method = cor_test)    
    tidy_res <- tidy(test)
    tidy_res
    if (plot == TRUE) {
    plt <-  ggplot(dat,  aes_string(x= col_1, y = col_2)) +
            geom_point() +
            labs(title = paste0(col_1, " compared to ", col_2),
                 subtitle = paste0("P-value is ", format.pval(tidy_res$p.value))) +
            theme_classic()
    return(plt)
    }
    return(tidy_res)
}

cogaps_quicK_check(joined_samp_e, "F.2", "Pattern_3", plot = T)

```
Top genes function
```{r}
dat <-joined_gene_e
pattern_n <- 3 
top_cogaps_genes <- function(dat, pattern_n, n_genes = 10, top_o_bot = "top"){
    if (top_o_bot == "top") {
        pattern <- sym(paste0("Pattern_",pattern_n))
        dat <- dat %>%
            arrange(desc({{pattern }} ))
    } else {
    pattern <- sym(paste0("Pattern_",pattern_n))
    dat <- dat %>%
        arrange({{pattern }})
    }
    
    dat <- dat %>% select(gene_id, gene_name, pattern)
    
    if (n_genes == "all") {
        return(dat)
    }
    
    return(head(dat, n_genes))
    
}

top_cogaps_genes(joined_gene_e, 3,n_genes = "all")
```

Find overlap of genes use patterns to find the highest number of overlap
```{r}
gene_cluster <- lung_clusts
cogaps_genes <- joined_gene_e
cluster_cogaps_sum <- function(gene_cluster, cogaps_genes){
    clust_only <- gene_cluster[,-1]
    data_out <- NULL
    for (cluster in seq(ncol(clust_only))) {
        genes <- unique(clust_only[,cluster])
        for (patterns in seq(ncol(cogaps_genes)- 2)) {
            cur_genes <- top_cogaps_genes(cogaps_genes,
                                          patterns, n_genes = "all") %>%
                filter(gene_name %in% genes)
            cur_res <- mean(cur_genes[,3])
            data_out <- rbind(data_out,
                c(colnames(clust_only)[cluster], paste0("Pattern_",patterns), cur_res))
        }
    }
    colnames(data_out) <- c("Clusters", "Patterns", "Mean Gene Association From Pattern")
    data_out <- as.data.frame(data_out) %>% arrange(desc(`Mean Gene Association From Pattern`))
    data_out[,3] <- as.numeric(data_out[,3])
    return(data_out)
    }


cluster_cogaps_sum(lung_clusts, joined_gene_e)

```
More Than 3 Clusters analysis
```{r}
gene_patterns_cogaps_u <- join_cogaps_gene_dat(cog_out_u, gtab_lung)
sample_patterns_cogaps_u <- join_cogaps_samp_dat(cog_out_u, lung_profiles)
top_clusters_per_pattern_u <- cluster_cogaps_sum(lung_clusts, gene_patterns_cogaps_u)

top_clusters_per_pattern_u
```

Function that use cluster_cogaps_sum output to plot matching patterns to profiles
```{r}
joined_samps <- join_cogaps_samp_dat(cog_out_u, lung_profiles)
cluster_gaps_sum_out <- top_clusters_per_pattern_u

matched_prof_pattern_plots <- function(joined_samps,
                                       cluster_gaps_sum_out,
                                       all_comparisons = TRUE,
                                       sig_plots = TRUE){
    plot_list <- list()
    df_out <- NULL
    ## Check last possible cluster
    ## Move down dataframe until all clusters have at least one selected top hit
    alphanum_clusters <- colnames(joined_samps)[!startsWith(colnames(joined_samps), "Pa") &
                                   !startsWith(colnames(joined_samps), "SA")]
    
    ## Make table of results via rbind
    for (cur_row in seq(nrow(cluster_gaps_sum_out))) {
        cur_row_full <- cluster_gaps_sum_out[cur_row,]
        
        new_row <- cogaps_quicK_check(joined_samps, cur_row_full[[1]],
                                      cur_row_full[[2]], plot = FALSE)
        new_row$comparison <- paste0(cur_row_full[[1]], "-vs-", cur_row_full[[2]])
        df_out <- rbind(df_out, new_row)
        
        }
    if (sig_plots == TRUE) {
        filt_df <- df_out %>% mutate(p.value_adj = p.adjust(p.value, method = "bonferroni")) %>%
            filter(p.value_adj < 0.05) %>%
            arrange(p.value_adj)
        for (row_filt in seq(nrow(filt_df))) {
            split_dat <- str_split(filt_df[row_filt,"comparison"],pattern = "-vs-")
            split_dat[[1]][2]
            plt <- cogaps_quicK_check(joined_samps, split_dat[[1]][1], split_dat[[1]][2], plot = T)
            plot_list[[filt_df[[row_filt,"comparison"]]]] <- plt
        }
        full_obj <- list()
        full_obj[["table"]] <- df_out
        full_obj[["plots"]] <- plot_list
        return(full_obj)
    }
    return(df_out)
}

test_out <-matched_prof_pattern_plots(sample_patterns_cogaps_u, top_clusters_per_pattern_u,sig_plots =  T)


test_out$plots

cogaps_quicK_check(joined_samp_e,  cur_row_full[[1]],  cur_row_full[[2]], plot = T)
cogaps_quicK_check(joined_samp_e, "F.2", "Pattern_3", plot = T)

test_out$table

## Pattern 11?
top_cogaps_genes(join_cogaps_gene_dat(cog_out_u, gtab_lung), pattern_n = 6,n_genes = 20)
```

Have a workable print out fo rall of these
```{r}
## Below lets us select a pattern
patterns_out_u <- patternMarkers(cog_out_u)
lapply(patterns_out_u$PatternMarkers, function(x){
    ind <- match(x, high_var_gtab$gene_id)
    renamed_dat <- high_var_gtab$gene_name[ind]
    
})

### Data frame output as follows
### Col 1 Cogaps patterns
### Col 2 best matching alphanum pattern
### Corr p-value between teh 2
### If cogaps could calculate a pattern marker
### List of cogaps genes or top gene if there is no true marker gene use my function
### genes from alphanum

cogs_obj <- cog_out_u
alphnum_clusts <- lung_clusts
alphnum_profiles <- lung_profiles
sample_patterns <- sample_patterns_cogaps_u
alphnum_cog_matched <- top_clusters_per_pattern_u
cogaps_genes <- cog_out_u
gtab <- high_var_gtab


cogs_obj = cog_out_d
alphnum_clusts = lung_clusts
alphnum_profiles = lung_profiles
gtab = high_var_gtab

cogaps_alphnum_comparison_df <- function(cogs_obj,
                                         alphnum_clusts,
                                         alphnum_profiles,
                                         gtab){
    ## Incorporate previous functions to make this require less functions
    
    sample_patterns <- join_cogaps_samp_dat(cogs_obj, lung_profiles)
    gene_patterns <- join_cogaps_gene_dat(cogs_obj, gtab)
    alphnum_cog_matched <- cluster_cogaps_sum(alphnum_clusts, gene_patterns)
    ## New code
    final_out <- NULL
    pval_df <- matched_prof_pattern_plots(sample_patterns, alphnum_cog_matched, sig_plots = F) %>%
        separate(comparison, c("alphnum", "cog_pattern"),sep = "-vs-", remove = FALSE)
    ## For loop for all patterns
    pattern_marks <- patternMarkers(cogs_obj)
    pattern_marks <- lapply(pattern_marks$PatternMarkers, function(x){
                        ind <- match(x, gtab$gene_id)
                        renamed_dat <- gtab$gene_name[ind]
    
                      })
    patterns <- names(pattern_marks)
    for (pattern in patterns) {
        #Col 1 is pattern
        col_1 <- pattern
        #Col 2 get best correlating alphnum
        top_res <- filter(pval_df, cog_pattern == pattern) %>%
            arrange(p.value) %>% slice(1)
        col_2 <- top_res$alphnum
        #Col 3 p value
        col_3 <- top_res$p.value
        # Col 4 pattern from cogaps or ours
        markers <- pattern_marks[[pattern]]
        
        col_4 <- ifelse(is_empty(markers), "cogaps_failed", "cogaps_worked")
        #Col 5 genes from patterns
        if (is_empty(markers)) {
            ### Pull top genes from alphnum data
            cogaps_genes <- join_cogaps_gene_dat(cogs_obj, gtab)
            alphnum_genes <- top_cogaps_genes(cogaps_genes,
                                             pattern_n = as.integer(stringr::str_extract(pattern, "\\d+")),
                                             n_genes = 10)
            ## In the future maybe come up with a filter on the top associated genes
            col_5 <- alphnum_genes$gene_name
             col_5 <- paste0(col_5, collapse = ", ")
        } else {
            col_5 <- pattern_marks[[pattern]]
            col_5 <- paste0(col_5, collapse = ", ")
        }
        ## Col 6 genes from most associatd alphnum clust
        col_6 <- alphnum_clusts[[top_res$alphnum]]
        col_6 <- col_6[col_6 != ""]   
        col_6 <- paste0(col_6, collapse = ", ")
    
            new_row <- as.data.frame(cbind(col_1, col_2, col_3, col_4, col_5, col_6))
            colnames(new_row) <- c("CoGAPS Pattern",
                                   "Alphnum Cluster",
                                   "P-value of sample correlation",
                                   "Can cogaps take out marker genes?",
                                   "Cogaps Pattern Markers",
                                   "Alphnum Cluster Genes")
            final_out <- rbind(final_out, new_row)
    }
    
    return(final_out)
    
}
```

```{r}


cog_analysis_d <- cogaps_alphnum_comparison_df(cogs_obj = cog_out_d,
                                     alphnum_clusts = lung_clusts,
                                     alphnum_profiles = lung_profiles,
                                     gtab = high_var_gtab)

cog_analysis_e <- cogaps_alphnum_comparison_df(cogs_obj = cog_out_e,
                                     alphnum_clusts = lung_clusts,
                                     alphnum_profiles = lung_profiles,
                                     gtab = high_var_gtab)

cog_analysis_u <- cogaps_alphnum_comparison_df(cogs_obj = cog_out_u,
                                     alphnum_clusts = lung_clusts,
                                     alphnum_profiles = lung_profiles,
                                     gtab = high_var_gtab)



# excel_out  <- list(paste0("coGAPS_patterns_",nrow(cog_analysis_d)) = cog_analysis_d,
#              paste0("coGAPS_patterns_",nrow(cog_analysis_e)) = cog_analysis_e,
#              paste0("coGAPS_patterns_",nrow(cog_analysis_u)) = cog_analysis_u,)

excel_out <- list(cog_analysis_d, cog_analysis_e, cog_analysis_u)
names(excel_out) <- c(paste0("coGAPS_patterns_",nrow(cog_analysis_d)),
                       paste0("coGAPS_patterns_",nrow(cog_analysis_e)),
                       paste0("coGAPS_patterns_",nrow(cog_analysis_u)))

openxlsx::write.xlsx(excel_out, file = paste0("data_out/3_cogaps_patterns_var_",cur_var,".xlsx"))
```

Save current environment
```{r}
#save.image("data_out/current_cogaps.rda")
save(cog_analysis_d, cog_analysis_e, cog_analysis_u,
     file = paste0("data_out/cogaps_var_",cur_var,""))
#load("data_out/current_cogaps.rda")
```

Load HPAStainR
```{r}
library(HPAStainR)
hpa_dat <- HPA_data_downloader(tissue_type = "both", save_file = FALSE)
HPAStainR(cog_analysis_u$`Cogaps Pattern Markers`[11],
          hpa_dat = hpa_dat$hpa_dat,cancer_dat = hpa_dat$cancer_dat,cancer_analysis = "normal")
```

```{r}
left_join(stab_lung,pheno_dat)

stab_lung$TRISCHD

pattern_clust_u <- join_cogaps_samp_dat(cog_out_u, lung_profiles)
joined_pattern_clust_stab <- left_join(pattern_clust_u, stab_lung)


joined_pattern_clust_stab$DTHVNT <- as.factor(joined_pattern_clust_stab$DTHVNT)

ggplot(joined_pattern_clust_stab, aes(DTHVNT, TRISCHD)) +
    geom_sina()

```

