---
title: "Cluster Report"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(viridisLite)
key_dat <- read.csv("../automated_cluster_analysis/data_out/key_table.csv")
```

# State of Clusters
The purpose of this Rmarkdown is to see how many clusters we have solved or at 
least have a hypothesis for.

# Current overview
```{r overview pre plot, echo=FALSE}
## Remove X clusters and prepare data for text
current_results <- read.csv("manual_curation.csv") %>%
    filter(!str_detect(label, "X"))
somewhat_solved <- sum(current_results$probable_cluster != "")
total_clusters <- nrow(current_results)
percent_solved <- round(somewhat_solved/total_clusters, 3) * 100

all_possibilites <- sort(table(current_results$probable_cluster),decreasing = T)
top_possibilites <- all_possibilites[all_possibilites >= 9]
current_results <- left_join(current_results, key_dat, by  = "label")

current_results <- mutate(current_results,
                          idea_what_it_is = ifelse(probable_cluster != "", "Yes", "No"),
                          approx_type = case_when(
                              probable_cluster == "SEX" ~ "Sex",
                              probable_cluster == "classic_contamination" ~ "Classic contam",
                              probable_cluster == "Immuno" ~ "Immunoglobs",
                              probable_cluster == "new_contamination" ~ "New contam",
                              probable_cluster == "ischemic" ~ "marked ischemic",
                              probable_cluster == "interleukins" ~ "marked interleukins",
                              TRUE ~ "Other"
                          ),
                          n_genes = str_count(genes, ";") + 1)
ggplot(current_results, aes(x = idea_what_it_is, fill = approx_type)) +
    geom_bar() +
    viridis::scale_fill_viridis(discrete = T) +
    theme_classic() +
    labs(title = "Overview of clusters we have hypotheses for",
         x="Hypothesis or not")

ggplot(current_results, aes(x = idea_what_it_is, fill = sex_sig)) +
    geom_bar() +
    viridis::scale_fill_viridis(discrete = T) +
    theme_classic() +
    labs(title = "Overview of clusters that correlate with sex",
         x="Hypothesis or not")

ggplot(current_results, aes(x = idea_what_it_is, fill = ischemic_sig)) +
    geom_bar() +
    viridis::scale_fill_viridis(discrete = T) +
    theme_classic() +
    labs(title = "Overview of clusters that correlate with ischemic time",
         x="Hypothesis or not")

ggplot(current_results, aes(x = n_genes, fill = idea_what_it_is)) +
    geom_histogram(alpha = 0.5, position = "identity", bins = 50) +
    theme_classic() +
    labs(title = "Distribution of solved clusters vs unsolved",
         x="N of genes in cluster")

ggplot(filter(current_results, n_genes < 75), aes(x = n_genes, fill = idea_what_it_is)) +
    geom_histogram(alpha = 0.5, position = "identity", bins = 40) +
    theme_classic() +
    labs(title = "Distribution of solved clusters vs unsolved; <70 genes",
         x="N of genes in cluster")


ggplot(current_results, aes(x = official_name, fill = idea_what_it_is)) +
    geom_bar() +
    viridis::scale_fill_viridis(discrete = T) +
    theme_classic() +
    coord_flip() +
    labs(title = "Solved clusters per tissue type",
         x="Tissue")


```

Currently we have `r total_clusters` clusters, with `r somewhat_solved` partially solved, meaning we have an idea on what they are. That means we currently have `r percent_solved`% clusters solved.


## Timeline of clusters solved
```{r pressure, echo=FALSE}
date_filename = "data_out/cluster_solved_history.csv"

if (!file.exists(date_filename)) {
    new_frame <- data.frame(somewhat_solved,Sys.Date())
    colnames(new_frame) <- c("solved_clusters", "date")
    write.csv(new_frame, date_filename, row.names = FALSE)
} else{
    date_frame <- read.csv(date_filename)
    date_frame$date <- as.Date(date_frame$date)
}

if (!(Sys.Date() %in% date_frame$date)) {
    new_row <- data.frame(somewhat_solved,Sys.Date())
    colnames(new_row) <- colnames(date_frame)
    date_frame <- rbind(date_frame, new_row)
    write.csv(date_frame, date_filename, row.names = FALSE)
}

ggplot(date_frame, aes(x = date, y = solved_clusters)) +
    geom_line() +
    geom_point() +
    theme_classic() +
    geom_hline(yintercept=nrow(current_results), color = "red") +
    ylim(0, nrow(current_results)) +
    labs(title = "N of clusters solved over time",
         y="N of solved clusters")


```
Load the excel file
```{r}
library(readxl)
excel_dat <- read_excel("../manual_cluster_analysis/manual_curation_11_09_21.xlsx")

table(excel_dat$generalized_cluster)

excel_filt <- excel_dat %>%
    mutate(generalized_cluster = case_when(
        generalized_cluster == "unsolved" ~ "unknown",
        generalized_cluster == "unknown" ~ "Unknown",
        generalized_cluster == "Immuno" ~ "Immunoglobulin",
        generalized_cluster == "new_contamination" ~ "Contamination",
        generalized_cluster == "classic_contamination" ~ "Contamination",
        generalized_cluster == "SELE cluster" ~ "Endothelial Activation",
        generalized_cluster == "Inflammation" ~ "Other",
        generalized_cluster == "Histones" ~ "Other",
        generalized_cluster == "Immune" ~ "Other",
        generalized_cluster == "Immuno/Mitochondria" ~ "Other",
        generalized_cluster == "FOS/FOSB" ~ "Other",
        T ~ generalized_cluster
    ),
    generalized_cluster = tools::toTitleCase(generalized_cluster)) %>%
    filter(generalized_cluster != "X CLUST")

excel_filt$generalized_cluster <- factor(excel_filt$generalized_cluster,
                                         levels = names(sort(table(excel_filt$generalized_cluster))))

table(excel_filt$generalized_cluster)

excel_filt %>%
    ggplot(., aes(y = generalized_cluster)) +
    geom_bar() +
    theme_bw() +
geom_text(stat='count', aes(label=..count..),nudge_x = 5)

counted_dat <- excel_filt %>% group_by(generalized_cluster) %>% count() %>%
    mutate(`Cluster Class` = case_when(
        generalized_cluster %in% c("SEX", "Contamination", "Accute Phase",
                                   "Immunoglobulin", "Endothelial Activation",
                                   "Mitochondria") ~ "Meta-Cluster",
       TRUE ~ "Tissue Specific/Mixed"
    ),
    n = ifelse(generalized_cluster == "SEX", n/2, n))

## Reorder data

require(viridisLite)
ggplot(counted_dat, aes(n, generalized_cluster, fill = `Cluster Class`)) +
    geom_bar(stat="identity") +
    xlim(0, 150) +
    geom_text(aes(label=n), hjust=-.05, size=3.5)+
    theme_minimal() +
    scale_fill_viridis(discrete = T) + 
    theme(legend.position = "top") +
    labs( y = "Cluster Type", x= "N of Clusters")
ggsave("data_out/cluster_type_fig.pdf",height = 5, width = 9)

ggsave("data_out/cluster_type_fig_thin.pdf",height = 5, width = 5)
```

Sex AGE and race plots
```{r}
phen_dat <- read.csv("../global_in/gtex_phenotypes_v8.csv")
subj_dat <- data.table::fread("../global_in/GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt")

sex_dat <- group_by(subj_dat, SEX) %>%
    summarize(sex_count = n()) %>%
    mutate(sex_percent = round(sex_count/nrow(subj_dat),digits = 2)) %>%
    mutate(SEX = ifelse(SEX == 1, "Male", "Female"))

sex_dat$sex_count <- c(636, 312)

age_dat <- group_by(subj_dat, AGE) %>%
    summarize(age_count = n()) %>%
    mutate(age_percent = round(age_count/nrow(subj_dat),digits = 2))

phen_dat %>%
    ggplot(., aes(AGE)) +
    geom_histogram(fill =  "white", color = "black") +
    theme_classic() +
    theme(text = element_text(size = 18))

ggplot(sex_dat, aes(sex_count, SEX, fill = SEX)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label=sex_count),nudge_x = -43, size=8)+
    theme_classic() +
    theme(text = element_text(size = 16)) +
    labs(x = "Count")
ggsave("sex_count.pdf", height = 3, width = 8 )
```

